### 2020-07-11 README.Rmd
fig 1 is ok

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# lithicprodswchinalatemiddlepleistocene

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/lithicprodswchinalatemiddlepleistocene/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): R Markdown source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `paper.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: data](/analysis/data): Data used in the analysis.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder:
    supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to run in your broswer or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

The simplest way to explore the text, code and data is to click on
[binder](https://mybinder.org/v2/gh/benmarwick/lithicprodswchinalatemiddlepleistocene/master?urlpath=rstudio)
to open an instance of RStudio in your browser, which will have the
compendium files ready to work with. Binder uses rocker-project.org
Docker images to ensure a consistent and reproducible computational
environment. These Docker images can also be used locally.

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.Rmd")` in the R console

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2020-07-11 analysis/paper/for paper ii.rmd
fig 1 is ok

---
output:
  word_document: default
  pdf_document: default
  html_document: default
---

Some normal text.

```{r message=FALSE, warning=FALSE, echo=FALSE}
#install required packages
library(dplyr)
library(knitr)
library(ggplot2)
library(reshape2)
library(readxl)
library(fitdistrplus)
library(plyr)
library(tidyr)
library(classInt)
library(knitr)
library(cowplot)

# read in the data

file_name <- here::here("analysis/data/raw_data/artefacts-after Sam_Dec2018_quina_modified.xls")

flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
debris <- as.data.frame(read_excel(file_name, sheet = "chunk&debris"))
retouch <- as.data.frame(read_excel(file_name, sheet = "retouch"))
core_flake_scars_1 <- as.data.frame(read_excel(file_name, sheet = "core flake scars"))
edge_angle_1 <- as.data.frame(read_excel(file_name, sheet = "edge angle"))
retouched_pebble_chunk <- as.data.frame(read_excel(file_name, sheet = "Pebble&chunk"))

scar_dir <- as.data.frame(read_excel(file_name, sheet = 'scar direction'))

core_flake_scars <- core_flake_scars_1[core_flake_scars_1$number %in% cores$number, ]

edge_angle <- edge_angle_1[edge_angle_1$number %in% retouch$number, ]

```

```{r message=FALSE, warning=FALSE, echo = FALSE}
# basic details of the dataset
artefact_ids <- list(flake_ids = flakes$number,
                     core_ids = cores$number,
                     debris_ids = debris$number)

retouch_ids <- retouch$number

# how many unique artefacts?

totals_of_each_type <- lapply(artefact_ids, function(i) length(i))

count_unique_artefacts <- sum(unlist(totals_of_each_type))

# debris pieces that are retouch pieces

debris_with_retouch <- intersect(retouch_ids, artefact_ids$debris_ids)

proportion_debris_retouched <- length(debris_with_retouch) / count_unique_artefacts

# proportion of each type

type_table <- data.frame(type = names(as.data.frame(totals_of_each_type)),
                         count = as.numeric(as.data.frame(totals_of_each_type)),
                         proportion = round(as.numeric(prop.table(as.data.frame(totals_of_each_type))),3))

# add total row at the bottom

type_table$type <- as.character(type_table$type)

type_table <- rbind(type_table, c("total", count_unique_artefacts, round(sum(type_table$proportion),2)))

# print pretty table

print("table of lithic types")

kable(type_table)

### raw materials

raw_material_table <- data.frame("category" = character(),
                                 'chert' = numeric(),
                                 'limestone' = numeric(),
                                 'basalt' = numeric(),
                                 'quartz' = numeric(),
                                 'sandstone' = numeric(),
                                 stringsAsFactors = FALSE)

row <- 1

raw_material_table[row, names(unlist(table(na.omit(flakes$material))))] <- 
  unlist(table(na.omit(flakes$material)))

raw_material_table[row, 1] <- "flake"
row <- row + 1

raw_material_table[row, names(unlist(table(na.omit(cores$material))))] <- 
  unlist(table(na.omit(cores$material)))
raw_material_table[row, 1] <- "core"
row <- row + 1

raw_material_table[row, names(unlist(table(na.omit(debris$material))))] <- 
  unlist(table(na.omit(debris$material)))
raw_material_table[row, 1] <- "debris"
row <- row + 1

raw_material_table[row, names(unlist(table(flakes$material[flakes$number %in% retouch$number])))] <- 
  unlist(table(flakes$material[flakes$number %in% retouch$number]))
raw_material_table[row, 1] <- "retouched pieces"
row <- row + 1

###tools raw materials

test <- as.matrix(t(table(flakes[flakes$number %in% retouch$number, 
                     c("material", "type2")])))

raw_material_table[row, colnames(test)] <- test["backed knife",]
raw_material_table[row, 1] <- "backed knife"
row <- row + 1

raw_material_table[row, colnames(test)] <- colSums( test[c("bec", "bec,scp"),])
raw_material_table[row, 1] <- "bec"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("borer","borer,dent", "borer,dent,scp","borer,end,scp","borer,notch",
                  "borer,notch,scp","borer,scp", "borer,scp,end","borer,strt" ), ])
raw_material_table[row, 1] <- "borer"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("BUR", "BUR? CORE?","BURIN" ), ])
raw_material_table[row, 1] <- "BUR"
row <- row + 1

raw_material_table[row, colnames(test)] <- test["chopper", ]
raw_material_table[row, 1] <- "chopper"
row <- row + 1

raw_material_table[row, colnames(test)] <- test["cleaver", ]
raw_material_table[row, 1] <- "cleaver"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("dent", "dent,borer,scp","dent,end","dent,notch",  "dent,scp",
                  "dent/NOT", "dent/POINT"  ), ])
raw_material_table[row, 1] <- "Denticular"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("end", "end,borer" , "end,borer,scp" ,"end,scp" ,"endscp" ), ])
raw_material_table[row, 1] <- "Endscraper"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("leva point","point" ,"POINT" , "point break" ,
                  "point,broken", "point?"  ), ])
raw_material_table[row, 1] <- "Point"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("moon","scp", "sidescp" , "sidescp-cvx"  ), ])
raw_material_table[row, 1] <- "scraper"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("NATBACK", "nature backed"), ])
raw_material_table[row, 1] <- "nature backed"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c( "notcch", "notch" , "notch,borer"  , "notch,end","notch,scp" ,       
                   "notch/DENT"), ])
raw_material_table[row, 1] <- "notch"
row <- row + 1

raw_material_table[row, colnames(test)] <- 
  colSums( test[c("TANG", "TANGED",  "TANGED POINT", "TANGED POINT /NOT", 
                  "tanged point?", "TANGED POINT?"  ), ])
raw_material_table[row, 1] <- "tanged point"
row <- row + 1

raw_material_table[is.na(raw_material_table)] <- 0 ##convert na to 0

raw_material_table$total <- rowSums(raw_material_table[,-1])

write.csv(raw_material_table, "raw_material_table.csv")

x <- raw_material_table

prop_raw_material <- cbind(category = x[, 1], round(x[, -1]/rowSums(x[, 2:(ncol(x)-1)])*100, digits = 1))

write.csv(prop_raw_material, "raw_material_table (percentage).csv")

print("tables for raw materials of the assemblage")

kable(raw_material_table)

print("proportion tables for raw materials of the assemblage")

kable(prop_raw_material)

### cores
core_summary <- cores[0, c(6:17)]

core_summary["mean",] <- c("mean" = apply(na.omit(cores[,c(6:17)]), 2, mean))

core_summary["sd",] <- c("sd" = apply(na.omit(cores[,c(6:17)]), 2, sd))

core_summary["CV",] <- core_summary["sd",]/core_summary["mean",]

core_summary <- rbind(core_summary, 
                      sapply(na.omit(cores[,6:17]), function(x) quantile(x, c(0.25, 0.5, 0.75))))

print("summary of core attributes")

kable(round(core_summary, digits = 1))

```

We will try univariate clustering for flake mass:

```{r message=FALSE, warning=FALSE, echo = FALSE}
require(Ckmeans.1d.dp)
result <- Ckmeans.1d.dp(flakes$mass[!is.na(flakes$mass)])

print("flake grouping result according to mass")
plot(result)

n_clusters <- length(result$size)
n_clusters_30 <- length(result$size[result$size > 30])

# assign cluster ID to table
flakes_clustered <- 
flakes %>% 
  filter(!is.na(mass)) %>% 
  mutate(cluster = result$cluster) %>% 
  right_join(flakes) %>% 
  filter(cluster %in% 1:n_clusters_30)

```

```{r message=FALSE, warning=FALSE, echo=FALSE}
# clean data to help with analysis
flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")
flakes <- flakes[!flakes$material %in% c('sandstone', 'quartz'),]

# how many levallois?
unique(flakes$type1)

leva <- flakes %>% 
  filter(grepl("LVT|LVF|LVFB", type1)) %>%
  filter(!grepl("LVF_", type1))

non_leva <- flakes %>% 
  filter(!grepl("LVT|LVF|LVFB", type1))

flakes_L <-  
  mutate(flakes, L = ifelse(grepl("LVT|LVF|LVFB", type1), "L", "N"))

# are the levallois pieces bigger?

print("compare the mass of Leva and non-leva flakes")
ggplot(flakes_L, aes(x = L, 
                     y = mass)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.2) +
  scale_y_log10() + theme_bw() +theme(text = element_text(size = 20))

# is it significant?

print("t-test to see if it is statistically significant?")

t.test(data = flakes_L, mass ~ L)
# no

# explore variation in thickness in these groups
thick_plot <- ggplot(flakes_L) +
  xlab("Thickness (mm)") + ylab("Density") +
  geom_density(aes(`Thickness_25%max`), colour = "red", alpha = 0.3) +
  geom_density(aes(`Thickness_50%max`), colour = "green", alpha = 0.3) +
  geom_density(aes(`Thickness_75%max`), colour = "blue", alpha = 0.3) +
  xlab("Thickness (mm)") + theme_bw() +
   theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

print("variation of thickness for all flakes (red = 25%max_dim,green = 50%max,blue=75%max")

thick_plot

# explore variation in width in these groups
width_plot <- ggplot(flakes_L) +
  xlab("Width (mm)") + ylab("Density") +
  geom_density(aes(`Width_25%max`), colour = "red", alpha = 0.3) +
  geom_density(aes(`Width_50%max`), colour = "green", alpha = 0.3) +
  geom_density(aes(`Width_75%max`), colour = "blue", alpha = 0.3) +
  xlab("Width (mm)") + theme_bw() +
   theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

print("variation of width for all flakes (red = 25%max_dim,green = 50%max,blue=75%max")

width_plot

# by size class that we identified earlier

print("compare thickness of leva and non-leva (red = 25%max_dim, 
      green = 50%max, blue=75%max")

thick_plot +
  facet_wrap(~ L, ncol = 1)

# need a test of uniformity (peak width)

# check the CVs
CV <- function(the_vector){
  mean_ <- mean(the_vector, na.rm = TRUE)
  sd_ <- sd(the_vector, na.rm = TRUE)
  cv <- (sd_/mean_) * 100
  round(cv, 3)
}

cv_res <- flakes_L %>% 
  group_by(L) %>% 
  dplyr::summarise(cv_25_thick = CV(`Thickness_25%max`),
                   cv_50_thick = CV(`Thickness_50%max`),
                   cv_75_thick = CV(`Thickness_75%max`),
                   cv_25_width = CV(`Width_25%max`), 
                   cv_50_width = CV(`Width_50%max`),
                   cv_75_width = CV(`Width_75%max`), 
                   n = n())

print("summary of cv for thickness and width for leva and non-leva flakes")

cv_res

```

The following are results for cores

```{r message=FALSE, warning=FALSE, echo=FALSE}

# function to compute frequencies and make a plot
f <-  function(the_data, the_column) {
  data.frame(table(the_data[the_column])) %>% 
    ggplot(aes(x = reorder(Var1, Freq), y = Freq)) +
    geom_bar(stat = "identity") +
    ylab("count") +
    xlab(the_column) + 
    theme_bw() +
    theme(axis.text.x = element_text(size = 20, angle = 45, hjust = 1, vjust = 1),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25)
    )
}

# some plots of frequencies for cores

## geometry
print("summary for core geometry")

f(cores, "core_geometry")

core_geom_table <-  table(na.omit(cores$core_geometry))

round(prop.table(core_geom_table), digits = 4)*100

## platform number

print("summary for core platform")

f(cores, "platform")

core_plat_table <-  table(na.omit(cores$platform))

round(prop.table(core_plat_table), digits = 4)*100

### cortex proportion of cores

print("cortex proportion of cores")

ggplot(cores, aes(cortex_percentage)) +
  geom_histogram(binwidth = 10) + theme_bw() +
  xlab("Cortex proportion (%)") + 
  ylab("Number of cores") +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

core_cortex_table <-  table(na.omit(cores$cortex_percentage))

round(prop.table(core_cortex_table), digits = 4)*100

## platform type

print("summary for core platform preparation")

f(cores, "platform_preparation")

core_plat_type <-  table(na.omit(cores$platform_preparation))

round(prop.table(core_plat_type), digits = 4)*100

## platform rotation
print("summary for core rotations")

f(cores, "rotations")

ggplot(cores, aes(rotations)) +
  geom_bar() + theme_bw() +
  xlab("Number of rotation") + 
  ylab("Number of cores") +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

core_plat_rotat <-  table(na.omit(cores$rotations))

round(prop.table(core_plat_rotat), digits = 4)*100

## core reduction type
print("summary for core reduction type")

f(cores, "type")

core_type_table <- table(na.omit(cores$type))

round(prop.table(core_type_table),digits = 4)*100

## core scar number

print("summary of core scar number")

ggplot(cores, aes(scar_number)) +
  geom_bar() + theme_bw() + 
  xlab("Scar number on cores") + 
  ylab("Count") +
    theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

core_scar_table <- table(na.omit(cores$scar_number))

round(prop.table(core_scar_table), digits = 4)*100

## core flake scar length

print("summary for core flake scar length")

core_flake_scars_melt <- melt(core_flake_scars_1)

colnames(core_flake_scars_melt)[3] <- "scar_length"

ggplot(core_flake_scars_melt[core_flake_scars_melt$scar_length < 500, ], 
       aes(scar_length)) +
  geom_density() + theme_bw() +
  xlab("Length of scar on cores") + 
  ylab("Density") +
    theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

print("median of scar length")

median(na.omit(core_flake_scars_melt[core_flake_scars_melt$scar_length < 500, ]$scar_length))

```

The following results are for flakes

```{r message=FALSE, warning=FALSE, echo=FALSE}

# explore flakes

flake_summary <- flakes[0, c(7:17, 19,20,22,23)]

flake_summary["mean",] <- c("mean" = apply(na.omit(flakes[,c(7:17, 19,20,22,23)]), 2, mean))

flake_summary["sd",] <- c("sd" = apply(na.omit(flakes[,c(7:17, 19,20,22,23)]), 2, sd))

flake_summary["cv", ] <- flake_summary["sd",] /flake_summary["mean",]

flake_summary <- rbind(flake_summary, 
                      sapply(na.omit(flakes[,c(7:17, 19,20,22,23)]), 
                             function(x) quantile(x, c(0.25, 0.5, 0.75))))

print("table for flake attributes")

kable(round(flake_summary, digits = 1))

## how many flake type1

print("summary for flake type1")
f(flakes, "type1")
table(na.omit(flakes$type1))
round(prop.table(table(na.omit(flakes$type1))), digits = 3)*100

## flake type2 
print("summary for flake type2")
f(flakes, "type2")
table(na.omit(flakes$type2))
round(prop.table(table(na.omit(flakes$type2))),digits = 3)*100

## how many retouched and unretouched flakes
flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")

flakes <- flakes[!flakes$material %in% c('qutz', 'quartz', 'basalt'),]

print("summary for flake retouch")

f(flakes, "retouched")
table(na.omit(flakes$retouched))
round(prop.table(table(na.omit(flakes$retouched))), digits = 3)*100

## raw material of flakes
print("summary for flake material")
f(flakes, "material")

## platform shape of flakes
print("summary for flake platform shape")
f(flakes, "platform_shape")

table(na.omit(flakes$platform_shape))
round(prop.table(table(na.omit(flakes$platform_shape))), digits = 4)*100

### number of flakes with distinctive platform
## platform of flakes
print("summary for flake platform")
f(flakes, "platform")

paste( "number of flakes with distinctive platform =", length(na.omit(flakes$platform)) )

table(na.omit(flakes$platform))
round(prop.table(table(na.omit(flakes$platform))), digits = 3)*100

# mass for all flakes
print("mass of flakes")
ggplot(flakes, aes(mass)) +
  geom_histogram(binwidth = 10) + theme_bw() +
  xlab("Mass (g)") + ylab("Number of flakes") +
    theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

## max dimension of all flakes

print("max dimension of all flakes")
ggplot(flakes, aes(max_dimension)) +
  geom_density() + theme_bw() +
  xlab("Maximum dimension (mm)") + ylab("Density") +
    theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

## length vs width at 50% max dime
print("length vs width at 50% max dim")

ggplot(flakes, aes(length, `Width_50%max`, colour = material)) +
  geom_point() +
  geom_smooth(method = "lm") + theme_bw() +
    theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

## platform shape vs thickness at 50% max dime
print("platform shape vs thickness at 50% max dim")

ggplot(flakes[!is.na(flakes$platform_shape),], 
       aes(`Thickness_50%max`, x = platform_shape)) +
  geom_boxplot() + scale_y_log10()+ theme_bw() +
  xlab("Platform shape") + ylab("Thickness at 50% max. dim. (mm)") +
    theme(axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust = 1),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

## platform shape vs max dim
print("platform shape vs max dim")

ggplot(flakes[!is.na(flakes$platform_shape),],
       aes(`max_dimension`, x = platform_shape)) +
  geom_boxplot() + scale_y_log10()+ theme_bw() +
   xlab("Platform shape") + ylab("Maximum dimension (mm)") +
    theme(axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust = 1),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

# mass for all flakes, retouched or not

print("mass for retouched and non-retouched flakes")

ggplot(flakes, aes(mass, colour = retouched)) +
  geom_density()+ theme_bw() +

# mass for all flakes, by raw material

print("mass for flakes of different raw materials")
ggplot(flakes, aes(mass, colour = material)) + scale_x_log10() +
  geom_density()+ theme_bw() +theme(text = element_text(size = 20))

## max dimension for all flakes

print("max dim for all flakes")
ggplot(flakes, aes(max_dimension)) +
  geom_histogram(binwidth = 5) + theme_bw() +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

## max dim for all flakes, retouched or not

print("max dim for all flakes, retouched or not")
ggplot(flakes, aes(`max_dimension`, colour = retouched)) +
  geom_density()+ theme_bw() +
  xlab("Maximum dimension (mm)") + ylab("Density") +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25),
        legend.text = element_text(size = 20))+
   theme(legend.position = c(0.8, 0.8),
         legend.title = element_blank())

## max dim for all flakes, by raw material
print("max dim for all flakes, by raw material")
ggplot(flakes, aes(`max_dimension`, colour = material)) +
  geom_density()+ theme_bw() +theme(text = element_text(size = 20))

## scar counts for all flakes, by raw material
print("scar counts for all flakes, by raw material")
ggplot(flakes, aes(`scar_number`)) +
  geom_bar()+ theme_bw() +
  xlab("Number of scars") + ylab("Count of flakes")+
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

flakes$scar_group <- ifelse(flakes$`scar_number` > 5, "more_than_5", "less_than_5")

ggplot(flakes, aes(mass, x = scar_group)) +
  geom_boxplot() +
  scale_y_log10()+ theme_bw() +theme(text = element_text(size = 20))

ggplot(flakes[!is.na(flakes$scar_group), ], aes(mass, colour = scar_group)) +
  geom_density()+ theme_bw() + scale_x_log10() +
  xlab("Mass (g)") + ylab("Count of flakes")+
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

# facetted plaforms

print("facetted platforms")

flakes$facetted_platform <- ifelse(grepl("fac", flakes$platform), "fac", "no")

table(na.omit(flakes$facetted_platform))
round(prop.table(table(na.omit(flakes$facetted_platform))), digits = 4)*100

ggplot(flakes, aes(`max_dimension`, colour = facetted_platform)) +
  geom_density()+ theme_bw() +
  xlab("Maximum dimension (mm)") + ylab("Density") +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25),
        legend.text = element_text(size = 20))+
   theme(legend.position = c(0.8, 0.8),
         legend.title = element_blank())

### cortex proportion of flakes

print("cortex proportion of flakes")

ggplot(flakes, aes(cortex_percentage)) +
  geom_histogram(binwidth = 10) + theme_bw() +
    xlab("Cortex proportion (%)") + ylab("Count of flakes") +
  theme(axis.text.x = element_text(size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 25),
          axis.title.y = element_text(size = 25))

### compare cortex for ret and non-ret

print("compare cortex for ret and non-retouched flakes")

ggplot(flakes, aes(x = cortex_percentage, fill = retouched)) + geom_histogram(binwidth = 10, position='dodge') + theme_bw() +theme(text = element_text(size = 20))

## how many points

print("how many points?")

flakes$point <- ifelse(grepl("point|POINT", flakes$type2), "point", 
                       ifelse(grepl("back", flakes$type2), "backed", "other"))

table(na.omit(flakes$point))
round(prop.table(table(na.omit(flakes$point))), digits = 4)*100

```

Group the flakes according to their

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Look to see if groups exist

# by flake size
plain_flakes <-  flakes[grepl("fl", flakes$type1), ]
d <- density(na.omit(plain_flakes$mass)) # returns the density data
# plot
plot(d)
ggplot(plain_flakes, aes(mass)) +
  geom_density()+ theme_bw() +theme(text = element_text(size = 20))

# test is we can discover natural breaks that divide up the flakes?

# how many groups should we look for?
n <- 5
natural_breaks <- classIntervals(plain_flakes$mass, n = n, 
                                 style = "jenks", dataPrecision = 2)
# plot the breaks
ggplot(plain_flakes, aes(mass)) +
  geom_density(colour = "blue") +
  geom_vline(xintercept = natural_breaks$brks, colour = "red")+ theme_bw() +theme(text = element_text(size = 20))

# create a df
plain_flakes$mass_group <- cut(plain_flakes$mass, 
                               breaks = c(0, natural_breaks$brks[-1]),
                               labels = 1:(length(natural_breaks$brks)-1))
# remove flakes not in a group
plain_flakes <- plain_flakes[!is.na(plain_flakes$mass_group), ]

# plot the groups
ggplot(plain_flakes, aes(mass_group, mass)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.1) +
  theme_bw() +theme(text = element_text(size = 20))

```

```{r message=FALSE, warning=FALSE, echo=FALSE}
# explore variation in thickness in these groups
thick_plot <- ggplot(plain_flakes) +
  xlab("Thickness (mm)") + ylab("Density") +
  geom_density(aes(`Thickness_25%max`), colour = "red", alpha = 0.3) +
  geom_density(aes(`Thickness_50%max`), colour = "green", alpha = 0.3) +
  geom_density(aes(`Thickness_75%max`), colour = "blue", alpha = 0.3) +
  xlab("Thickness (mm)") + theme_bw() +theme(text = element_text(size = 20))

thick_plot

# by size class that we identified earlier
thick_plot +
  facet_wrap(~ mass_group, ncol = 1)

# Lycett and Eren found out the thickness is more evenly distributed and less variable across preferential Levallois flakes. So , let's get the CV values for thickness for each group
CV <- function(the_vector){
  mean_ <- mean(the_vector, na.rm = TRUE)
  sd_ <- sd(the_vector, na.rm = TRUE)
  cv <- (sd_/mean_)*100
  round(cv, 3)
}

plain_flakes %>% 
  group_by(mass_group) %>% 
  dplyr::summarise(cv_25_thick = CV(`Thickness_25%max`),
                   cv_50_thick = CV(`Thickness_50%max`),
                   cv_75_thick = CV(`Thickness_75%max`),
                   cv_25_width = CV(`Width_25%max`), 
                   cv_50_width = CV(`Width_50%max`),
                   cv_75_width = CV(`Width_75%max`), 
                   n = n())

```

Results for edge angles

```{r message=FALSE, warning=FALSE, echo=FALSE}
# calculate the edge angle

edge_angle_2 <- edge_angle

edge_angle_2[, 2:(ncol(edge_angle_2)-1)] <- atan(edge_angle[, 2:(ncol(edge_angle_2)-1)]/2/3)/pi*180*2

df_angle <- melt(edge_angle_2[, 1:(ncol(edge_angle_2)-1)])

df_angle$variable <- substr(df_angle$variable, 1, 9)

colnames(df_angle)[2:3] <- c("section", "angle")

# plot the angles

print("edge angels")

ggplot(df_angle, aes(x = angle, col = section)) + geom_density() + theme_bw() +theme(text = element_text(size = 20))

## find quina and compare their angles.

quina_id <- flakes$number[which(flakes$type3 == "quina")]

df_angle$quina <- ifelse(df_angle$number %in% quina_id, "quina", "no")

ggplot(df_angle, aes(x = angle, col = quina)) + geom_density() + theme_bw() +theme(text = element_text(size = 20))

ggplot(df_angle, aes(x = angle, col = section)) + geom_density() + theme_bw() +theme(text = element_text(size = 20)) + facet_wrap(~quina)

## compare thickness of quina and other

df_thick <- melt(flakes[, c("number", "Thickness_25%max", "Thickness_50%max", "Thickness_75%max")])

df_thick$quina <- ifelse(df_thick$number %in% quina_id, "quina", "no")

ggplot(df_thick, aes(x = value, col = quina)) + geom_density() + theme_bw() +theme(text = element_text(size = 20)) + facet_wrap(~variable)

# plot the thickness of 50% max

print("thickness of 50% max dim")
ggplot(df_thick[df_thick$variable =="Thickness_50%max", ], aes(quina, value)) +
  geom_boxplot() + theme_bw() +theme(text = element_text(size = 20))

## count the number of shapes of edges.

edge_shapes <- unlist(strsplit(retouch$`edge shape`, ","))

data.frame(table(edge_shapes)) %>% 
ggplot(aes(x = reorder(edge_shapes, Freq), y = Freq)) +
    geom_bar(stat = "identity") +
    ylab("n") +
    xlab("Edge type") + 
    theme_bw() +theme(text = element_text(size = 20)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

data.frame(table(edge_shapes))

## number of layers

f(retouch, 'number of layers')

## number of edges

f(retouch, 'number of edge')

## number of layers for each number of eadge

df_layer_edge <- melt(retouch[, 1:3])

 data.frame(table(retouch[,2:3])) %>% 
    ggplot(aes(x = reorder(number.of.layers, Freq), y = Freq)) +
    geom_bar(stat = "identity") +
    ylab("n") +
    xlab("number of layers") + 
    theme_bw() +theme(text = element_text(size = 20)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    facet_wrap(~number.of.edge)

```

scar number and direction

```{r message=FALSE, warning=FALSE, echo=FALSE}

# scar number and direction

## count how many scars from each direction

print("count how many scars from each direction")

table(unlist(scar_dir[,2:8]))

## How many % flakes that have dorsal scars from at least 3 complete different directions

print("How many % flakes that have dorsal scars from at least 3 complete different directions")

table(unlist(scar_dir[,2:8]))

## calculate GIUR

df_giur <- retouch[, c(1,8,10,12,14,16,18,20,22)]

df_giur[, -1] <- df_giur[, -1] / retouch[, c(9,11,13,15,17,19,21,23)]

df_giur_clustered <- left_join(df_giur, flakes_clustered[, c("number", "cluster")])

df_giur_clustered_melt <- melt(df_giur_clustered[,-1], id.vars = "cluster")

## compare GIUR with edge angle

df_giur_edge <- melt(df_giur)

df_giur_edge$variable <- substr(df_giur_edge$variable, 1, 9)

colnames(df_giur_edge)[2:3] <- c("section", "giur")

df_giur_edge <- left_join(df_giur_edge[df_giur_edge$giur <= 1, ], df_angle)

print("compare the GIUR and edge angle for different sections")

ggplot(na.omit(df_giur_edge), aes(angle, giur, colour = section)) +
  geom_point() +
  geom_smooth(method = "lm") + theme_bw() +theme(text = element_text(size = 20))

print("compare the GIUR and edge angle for all sections")

ggplot(na.omit(df_giur_edge), aes(angle, giur)) +
  geom_point() +
  geom_smooth(method = "lm") + theme_bw() +theme(text = element_text(size = 20))

## compare GIUR for different mass categories

print("compare GIUR for different mass groups")

ggplot(na.omit(df_giur_clustered_melt[df_giur_clustered_melt$value <=1,]), aes(value)) +
  geom_histogram() + theme_bw() +theme(text = element_text(size = 20)) + facet_wrap(~cluster)

## calculate the median GIUR for each group

print("the median GIUR for each group")

tapply(df_giur_clustered_melt$value, df_giur_clustered_melt$cluster, median, na.rm=TRUE)

## compare GIUR for Quina and non-quina

print("Compare GIUR for Quina and non-quina")

df_giur_clustered$quina <- ifelse(df_giur_clustered$number %in% quina_id, "quina", "no")

df_giur_clustered_melt_2 <- melt(df_giur_clustered[,-1], id.vars = "quina")

ggplot(na.omit(df_giur_clustered_melt_2[df_giur_clustered_melt_2$value <=1,]), aes(value)) +
  geom_histogram() + theme_bw() +theme(text = element_text(size = 20)) + facet_wrap(~quina)

print("Histogram of GIUR for Quina")

ggplot(na.omit(df_giur_clustered_melt_2[df_giur_clustered_melt_2$value <=1 &
                                          df_giur_clustered_melt_2$quina == "quina",]), aes(value)) +
  geom_histogram()  + theme_bw() +theme(text = element_text(size = 20))

print("Histogram of GIUR for non Quina")

ggplot(na.omit(df_giur_clustered_melt_2[df_giur_clustered_melt_2$value <=1 &
                                          df_giur_clustered_melt_2$quina == "no",]), aes(value)) +
  geom_histogram() +  theme_bw() +theme(text = element_text(size = 20))

print("the median GIUR for quina and non-quina")

tapply(df_giur_clustered_melt_2$value, df_giur_clustered_melt_2$quina, median, na.rm=TRUE)

# Invasiveness index
print("results for invasiveness index")

df_invasive <- retouch[, c(1, 24:39)]

df_invasive_clustered <- left_join(df_invasive, flakes_clustered[, c("number", "cluster")])

df_invasive_clustered_melt <- melt(df_invasive_clustered[,-1], id.vars = "cluster")

invasive_table <- as.data.frame(table(df_invasive_clustered_melt[, c("cluster", "value")]))

ggplot(invasive_table[invasive_table$value %in% c(0.5, 1.0), ], aes(fill=value, y= Freq, x=cluster)) + 
    geom_bar( stat="identity", position="fill", width = 0.6) + theme_bw() +theme(text = element_text(size = 20)) +
    theme(axis.text=element_text(size=14), axis.title=element_text(size=16,face="bold")) +
    theme(legend.text=element_text(size=14))

```

### 2020-07-11 analysis/paper/fig-1-compare-levallois-non-levallois.R

@@ -1,0 +1,261 @@
+
+
+library(dplyr)
+library(knitr)
+library(ggplot2)
+library(reshape2)
+library(readxl)
+library(fitdistrplus)
+library(plyr)
+library(tidyr)
+library(ggbeeswarm)
+library(knitr)
+library(cowplot)
+
+# read in the data --------------------------------------------------
+
+file_name <- here::here("analysis/data/raw_data/artefacts-after Sam_Dec2018_quina_modified.xls")
+flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
+cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
+debris <- as.data.frame(read_excel(file_name, sheet = "chunk&debris"))
+retouch <- as.data.frame(read_excel(file_name, sheet = "retouch"))
+core_flake_scars_1 <- as.data.frame(read_excel(file_name, sheet = "core flake scars"))
+edge_angle_1 <- as.data.frame(read_excel(file_name, sheet = "edge angle"))
+retouched_pebble_chunk <- as.data.frame(read_excel(file_name, sheet = "Pebble&chunk"))
+scar_dir <- as.data.frame(read_excel(file_name, sheet = 'scar direction'))
+core_flake_scars <- core_flake_scars_1[core_flake_scars_1$number %in% cores$number, ]
+edge_angle <- edge_angle_1[edge_angle_1$number %in% retouch$number, ]
+
+# prepare the data --------------------------------------------------
+
+require(Ckmeans.1d.dp)
+result <- Ckmeans.1d.dp(flakes$mass[!is.na(flakes$mass)])
+
+n_clusters <- length(result$size)
+n_clusters_30 <- length(result$size[result$size > 30])
+
+# assign cluster ID to table
+flakes_clustered <-
+  flakes %>%
+  filter(!is.na(mass)) %>%
+  mutate(cluster = result$cluster) %>%
+  right_join(flakes) %>%
+  filter(cluster %in% 1:n_clusters_30)
+
+# clean data to help with analysis
+flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")
+flakes <- flakes[!flakes$material %in% c('sandstone', 'quartz'),]
+
+leva <- flakes %>%
+  filter(grepl("LVT|LVF|LVFB", type1)) %>%
+  filter(!grepl("LVF_", type1))
+
[…]

### 2020-07-11 analysis/paper/script for paper II.R

@@ -1,0 +1,526 @@
+
+#install required packages
+library(dplyr)
+library(knitr)
+library(ggplot2)
+library(readxl)
+library(fitdistrplus)
+library(plyr)
+library(tidyr)
+library(classInt)
+
+
+
+
+# read in the data
+
+file_name <- "artefacts-after Sam_Dec2018_quina_modified.xls" 
+
+flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
+cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
+debris <- as.data.frame(read_excel(file_name, sheet = "chunk&debris"))
+retouch_1 <- as.data.frame(read_excel(file_name, sheet = "retouch"))
+core_flake_scars_1 <- as.data.frame(read_excel(file_name, sheet = "core flake scars"))
+edge_angle_1 <- as.data.frame(read_excel(file_name, sheet = "edge angle"))
+retouched_pebble_chunk <- as.data.frame(read_excel(file_name, sheet = "Pebble&chunk"))
+
+
+### raw materials
+
+raw_material_table <- data.frame("category" = character(),
+                                 'chert' = numeric(),
+                                 'limestone' = numeric(),
+                                 'basalt' = numeric(),
+                                 'quartz' = numeric(),
+                                 'sandstone' = numeric(),
+                                 stringsAsFactors = FALSE)
+
+row <- 1
+
+raw_material_table[row, names(unlist(table(na.omit(flakes$material))))] <- 
+  unlist(table(na.omit(flakes$material)))
+
+raw_material_table[row, 1] <- "flake"
+row <- row + 1
+
+
+raw_material_table[row, names(unlist(table(na.omit(cores$material))))] <- 
+  unlist(table(na.omit(cores$material)))
+raw_material_table[row, 1] <- "core"
+row <- row + 1
+
+raw_material_table[row, names(unlist(table(na.omit(debris$material))))] <- 
+  unlist(table(na.omit(debris$material)))
+raw_material_table[row, 1] <- "debris"
+row <- row + 1
+
+
+raw_material_table[row, names(unlist(table(flakes$material[flakes$number %in% retouch_1$number])))] <- 
+  unlist(table(flakes$material[flakes$number %in% retouch_1$number]))
[…]

### 2020-07-11 analysis/paper/script for paper II_2.R

@@ -1,0 +1,1351 @@
+
+#install required packages
+library(dplyr)
+library(knitr)
+library(ggplot2)
+library(reshape2)
+library(readxl)
+library(fitdistrplus)
+library(tidyr)
+library(classInt)
+library(cowplot)
+library(gridExtra)
+
+# read in the data
+
+
+file_name <- here::here("analysis/data/raw_data/artefacts-after Sam_May2020_quina_modified.xls")
+
+flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
+cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
+debris <- as.data.frame(read_excel(file_name, sheet = "chunk&debris"))
+retouch <- as.data.frame(read_excel(file_name, sheet = "retouch"))
+core_flake_scars_1 <- as.data.frame(read_excel(file_name, sheet = "core flake scars"))
+edge_angle_1 <- as.data.frame(read_excel(file_name, sheet = "edge angle"))
+retouched_pebble_chunk <- as.data.frame(read_excel(file_name, sheet = "Pebble&chunk"))
+
+scar_dir <- as.data.frame(read_excel(file_name, sheet = 'scar direction'))
+
+core_flake_scars <- core_flake_scars_1[core_flake_scars_1$number %in% cores$number, ]
+
+edge_angle <- edge_angle_1[edge_angle_1$number %in% retouch$number, ]
+
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+# basic details of the dataset
+
+artefact_ids <- list(flake_ids = flakes$number,
+                     core_ids = cores$number,
+                     debris_ids = debris$number)
+
+retouch_ids <- retouch$number
+
+# how many unique artefacts?
+
+totals_of_each_type <- lapply(artefact_ids, function(i) length(i))
+
+count_unique_artefacts <- sum(unlist(totals_of_each_type))
+
+# debris pieces that are retouch pieces
+
+debris_with_retouch <- intersect(retouch_ids, artefact_ids$debris_ids)
+
+proportion_debris_retouched <- length(debris_with_retouch) / count_unique_artefacts
+
+# proportion of each type
+
+type_table <- data.frame(type = names(as.data.frame(totals_of_each_type)),
+                         count = as.numeric(as.data.frame(totals_of_each_type)),
[…]

### 2020-07-11 analysis/paper/table-1-descriptive-stats-levallois-non-levallois-flakes.R

@@ -1,0 +1,103 @@
+
+
+#install required packages
+library(dplyr)
+library(knitr)
+library(ggplot2)
+library(reshape2)
+library(readxl)
+library(fitdistrplus)
+library(plyr)
+library(tidyr)
+library(classInt)
+library(knitr)
+library(cowplot)
+
+# read in the data
+
+
+file_name <- here::here("analysis/data/raw_data/artefacts-after Sam_Dec2018_quina_modified.xls")
+
+flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
+cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
+debris <- as.data.frame(read_excel(file_name, sheet = "chunk&debris"))
+retouch <- as.data.frame(read_excel(file_name, sheet = "retouch"))
+core_flake_scars_1 <- as.data.frame(read_excel(file_name, sheet = "core flake scars"))
+edge_angle_1 <- as.data.frame(read_excel(file_name, sheet = "edge angle"))
+retouched_pebble_chunk <- as.data.frame(read_excel(file_name, sheet = "Pebble&chunk"))
+scar_dir <- as.data.frame(read_excel(file_name, sheet = 'scar direction'))
+core_flake_scars <- core_flake_scars_1[core_flake_scars_1$number %in% cores$number, ]
+edge_angle <- edge_angle_1[edge_angle_1$number %in% retouch$number, ]
+
+
+# basic details of the dataset
+artefact_ids <- list(flake_ids = flakes$number,
+                     core_ids = cores$number,
+                     debris_ids = debris$number)
+
+retouch_ids <- retouch$number
+
+# clean data to help with analysis
+flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")
+flakes <- flakes[!flakes$material %in% c('sandstone', 'quartz'),]
+
+leva <- flakes %>%
+  filter(grepl("LVT|LVF|LVFB", type1)) %>%
+  filter(!grepl("LVF_", type1))
+
+non_leva <- flakes %>%
+  filter(!grepl("LVT|LVF|LVFB", type1))
+
+flakes_L <-
+  mutate(flakes, L = ifelse(grepl("LVT|LVF|LVFB", type1), "L", "N"))
+
+
+# need a test of uniformity (peak width)
+
+# check the CVs
[…]

### 2020-10-04 analysis/paper/old-stuff/flake_attributes_and_size.Rmd
tidy up and polish fig 1

---
title: "Guanyingdong Stone Artefact Assemblage Report: Variation by Size"
author: "HY and BM"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   bookdown::word_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
---

```{r setup-flake-size, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Nature’s standard figure sizes are 89 mm (single column) and 183 mm (double column)
 # https://www.nature.com/nature/for-authors/formatting-guide
```

THe reviewers ask:

>  can they offer other data that might elucidate changes in flaking type with reduction, e.g. changes in flake attributes that correlate with size declines?

Read in the data, deal with dupicates, get the basic flake details, run to L454 of hu_artefacts_report.Rmd, then continue from here.

Let's see if there are any natural groupings in flake mass:

```{r}

flake_hist_1 <- 
ggplot(flakes,
       aes(mass)) +
  geom_histogram() +
  theme_minimal() +
  xlab("Mass (g)")

flake_hist_2 <- 
ggplot(flakes,
       aes(mass)) +
  geom_histogram() +
  theme_minimal() +
  scale_x_log10() +
  xlab("Mass (g)")

library(cowplot)

flake_mass_distributions <- plot_grid(flake_hist_1, flake_hist_2, ncol = 1)

```

Nothing obvious, let's try a little clustering. We will try univariate clustering:

```{r}
require(Ckmeans.1d.dp)
result <- Ckmeans.1d.dp(flakes$mass[!is.na(flakes$mass)])
plot(result)

n_clusters <- length(result$size)
n_clusters_30 <- length(result$size[result$size > 30])

# assign cluster ID to table
flakes_clustered <- 
flakes %>% 
  filter(!is.na(mass)) %>% 
  mutate(cluster = result$cluster) %>% 
  right_join(flakes) %>% 
  filter(cluster %in% 1:n_clusters_30)
```

We get `r n_clusters` clusters, with `r n_clusters_30` that have more than 30 artefacts in them

```{r}
ggplot(flakes_clustered,
       aes(mass, 
          length,
          colour = factor(cluster)))+
  geom_point()

flake_hist_with_clusters <- 
ggplot(flakes_clustered,
       aes(mass)) +
  geom_histogram(aes(fill = factor(cluster))) +
  annotate("text", x = 240, y = 120, label = glue::glue('Total number of artefacts in sample: {nrow(flakes_clustered)}')) +
  scale_fill_viridis_d(name = "Size class") +
  theme_minimal() +
  xlab("Mass (g)")  +
  coord_equal(ratio = 0.2) +
  theme(legend.position = c(.75, .75),
        legend.direction = "horizontal",
        legend.key.size = unit(0.6,"line"),
        legend.key.width = unit(0.6,"line"),
        legend.key.height = unit(0.6,"line"),
        legend.text = element_text(size=8),
        legend.box.background = element_rect(fill = "white", colour = "white"))
```

Now let's see how some of the key variables change accross these size clusters:

We see that some dimensions don't increase with size. Platform thickness decreases a little, as mass goes up, platform width becomes more variable. Scar numbers have less high outliers as size increases.

```{r}

library(stringr)
# get sample sizes
flakes_clustered_numeric_vars_sample_sizes <- 
flakes_clustered %>% 
  select_if(is.numeric) %>%
  gather(variable, value, -cluster) %>% 
  mutate(variable = str_to_title(str_replace_all(variable, "_", " "))) %>% 
  filter(!is.na(value)) %>% 
  mutate(variable = str_replace_all(variable, "%", "% ")) %>% 
  mutate(cluster = factor(cluster)) %>% 
  group_by(variable) %>% 
  tally()

# add measurement units
flakes_clustered_numeric_vars_sample_sizes$units <- 
  c("", 
    "mm, ",
    "g, ",
    "mm, ",
    "mm, ",
    "mm, ",
    "mm, ",
    "mm, ",
    "",
    "mm, ",
    "mm, ",
    "mm, ",
    "mm, ",
    "mm, ",
    "mm, ")

flakes_clustered_numeric_vars_plot <- 
flakes_clustered %>% 
  select_if(is.numeric) %>%
  gather(variable, value, -cluster) %>% 
  mutate(variable = str_to_title(str_replace_all(variable, "_", " "))) %>% 
  filter(!is.na(value)) %>% 
  mutate(variable = str_replace_all(variable, "%", "% ")) %>% 
  mutate(cluster = factor(cluster)) %>% 
  left_join(flakes_clustered_numeric_vars_sample_sizes) %>% 
  # mutate(variable = glue::glue('{variable} ({units}n = {n})')) %>% 
  ggplot(aes(cluster,
             value)) +
  geom_boxplot() +
  facet_wrap( ~ variable, scales = "free_y") +
  xlab("Size class") +
  ylab("") +
  theme_minimal(base_size = 8)

figure_one <- 
cowplot::plot_grid(flake_hist_with_clusters, 
                   flakes_clustered_numeric_vars_plot,
                   labels = "AUTO", 
                   ncol = 1, 
                   align = 'v', 
                   axis = 'lr', 
                   rel_heights = c(0.5, 1.5)) #

save_plot("Extended_Data_Figure_X.png", figure_one,
          base_aspect_ratio = 1, 
          base_height = 8,
          base_width = 18.3/2.5)
```

Platform thickness decreases, width at 50% and 75% does not increase as much as width at 25% suggesting a slight increase in triangular-shaped flakes over parallel-sided flakes as the flakes get larger.

How about the categorical variables?

```{r}

flakes_categorical <- 
flakes_clustered %>% 
  select_if(function(col) is.character(col) | 
                                   all(col == .$cluster)) %>% 
  gather(variable, value, -cluster) %>% 
  mutate(cluster = factor(cluster)) %>% 
  filter(!is.na(value)) %>% 
  group_by(cluster, variable, value) %>% 
  tally() %>% 
  ungroup() %>% 
  group_by(cluster, variable) %>%  
  mutate(prop = n / sum(n) * 100)

# tidy up some of these variables
library(stringr)
library(forcats)

flakes_clustered_cats <- 
flakes_clustered %>% 
  select_if(function(col) is.character(col) | 
                                   all(col == .$cluster))

# raw material -----------------------------------------------------------------
# material, ok
flakes_clustered_cats %>% 
 group_by(material) %>% 
 tally(sort = TRUE)

# make table for plotting
flakes_clustered_cats_raw_material <- 
flakes_clustered_cats %>%  
 group_by(`material` , cluster) %>% 
 tally(sort = TRUE)

# flake type -----------------------------------------------------------------

# type1, ok
flakes_clustered_cats %>% 
 group_by(type1) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_flake_types <- 
  flakes_clustered_cats %>% 
  mutate(type1 = if_else(str_detect(type1, "lvf"), "Levallois", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "ret"), "retouch", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "blade|bld"), "blade", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "flk|flake"), "flake", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "dbd"), "débordant", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "kbw"), "Kombewa", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "end"), "end scraper", type1)) %>% 
  mutate(type1 = fct_lump(type1, 5))

# check again
flakes_clustered_cats_flake_types %>% 
 group_by(type1) %>% 
 tally(sort = TRUE)

# for plotting
flakes_clustered_cats_flake_types <- 
flakes_clustered_cats_flake_types %>% 
 group_by(`type1`, cluster) %>% 
 tally(sort = TRUE)

# retouch type -----------------------------------------------------------------

# type 2, special case , retouch types
flakes_clustered_cats %>% 
 group_by(type2) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_retouch_type <- 
flakes_clustered_cats %>% 
 separate(type2, into = c("type2a", "type2b", "type2c"), sep = ",") %>% 
  mutate(type2a = if_else(str_detect(type2a, "lvf"), "levallois", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "ret"), "retouch", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "blade|bld"), "blade", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "flk|flake"), "flake", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "tang"), "tanged", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "not"), "notch", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "bur"), "burin", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "point"), "point", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "dent"), "denticulate", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "scp"), "scraper", type2a)) %>% 
  mutate(type2a = str_squish(type2a))  %>% 
  mutate(type2a = fct_lump(type2a, 7)) %>% 
  filter(!is.na(type2a))

# check it
flakes_clustered_cats_retouch_type %>% 
 group_by(type2a) %>% 
 tally(sort = TRUE)

# for plotting
flakes_clustered_cats_retouch_type <- 
flakes_clustered_cats_retouch_type %>% 
# rename(`Retouch type` = type2a ) %>% 
# group_by(`Retouch type`, cluster) %>% 
 tally(sort = TRUE)

# platform type -----------------------------------------------------------------

# Platform, ok
flakes_clustered_cats %>% 
 group_by(platform) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_platform_type  <- 
  flakes_clustered_cats %>% 
   mutate(platform = if_else(str_detect(platform, "facet"), "facetted", platform)) %>% 
   mutate(platform = if_else(str_detect(platform, "cotex|cortex"), "cortex", platform)) %>%
   mutate(platform = fct_lump(platform, 6)) %>% 
  filter(!is.na(platform))

# for plotting
flakes_clustered_cats_platform_type <- 
flakes_clustered_cats_platform_type %>% 
 group_by(`platform`, cluster) %>% 
 tally(sort = TRUE)

# platform shape -----------------------------------------------------------------

# platform shape, ok
flakes_clustered_cats %>% 
 group_by(platform_shape) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_platform_shape <- 
flakes_clustered_cats %>% 
  mutate(platform_shape = if_else(platform_shape == "sb", "quadrangle", platform_shape)) %>%
  mutate(platform_shape = if_else(platform_shape == "cdg", "chapeau de gendarme", platform_shape)) %>%
  mutate(platform_shape = if_else(platform_shape == "sx", "rhombus", platform_shape)) %>%
  mutate(platform_shape = fct_lump(platform_shape, 4))  %>% 
  filter(!is.na(platform_shape))

# for plotting
flakes_clustered_cats_platform_shape <- 
flakes_clustered_cats_platform_shape %>% 
# rename(`platform` = platform_shape ) %>% 
 group_by(`platform_shape`, cluster) %>% 
 tally(sort = TRUE)

# cortex texture-----------------------------------------------------------------

# cortex texture, ok 
flakes_clustered_cats %>% 
 group_by(cortex_texture) %>% 
 tally(sort = TRUE)

# for plotting
flakes_clustered_cats_cortex_texture <- 
flakes_clustered_cats %>% 
 group_by(`cortex_texture`, cluster) %>% 
 filter(!is.na(`cortex_texture` )) %>% 
 tally(sort = TRUE)

# cortex location-----------------------------------------------------------------

# cortex location, ok
flakes_clustered_cats %>% 
 group_by(cortex_location) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_cortex_location <- 
  flakes_clustered_cats %>% 
   mutate(cortex_location = if_else(str_detect(cortex_location, "prox"), 
                                    "proximal", 
                                    cortex_location)) %>%
   mutate(cortex_location = fct_lump(cortex_location, 6)) %>% 
  filter(!is.na(cortex_location),
         cortex_location != "vent")

flakes_clustered_cats_cortex_location <- 
flakes_clustered_cats_cortex_location %>% 
 group_by(`cortex_location`, cluster) %>% 
 filter(!is.na(`cortex_location` )) %>% 
 tally(sort = TRUE)

# cortex location-----------------------------------------------------------------

# initiation, ok 
flakes_clustered_cats %>% 
 group_by(initiation) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_initiation <- 
  flakes_clustered_cats %>% 
   mutate(initiation = fct_lump(initiation, 3))

flakes_clustered_cats_initiation <- 
flakes_clustered_cats_initiation %>% 
 group_by(`initiation`, cluster) %>% 
 filter(!is.na(`initiation` )) %>% 
 tally(sort = TRUE)

# cortex location-----------------------------------------------------------------

# platform preparation, ok, ignore, no meaningful info in here
flakes_clustered_cats %>% 
 group_by(platform_preparation) %>% 
 tally(sort = TRUE)

flakes_clustered_cats <- 
  flakes_clustered_cats %>% 
   mutate(platform_preparation = if_else(str_detect(platform_preparation, "fac"), "facetted", platform_preparation)) %>% 
   mutate(platform_preparation = fct_lump(platform_preparation, 3))

```

Make a nice panel plot, we have separate dataframes for each variable.

```{r}
list_of_dfs <- list(
  #flakes_clustered_cats_retouch_type,
                    flakes_clustered_cats_raw_material,
                    flakes_clustered_cats_platform_type,
                    flakes_clustered_cats_platform_shape,
                    flakes_clustered_cats_initiation,
                    flakes_clustered_cats_flake_types,
                    flakes_clustered_cats_cortex_texture,
                    flakes_clustered_cats_cortex_location)

library(purrr)

map_dfr(list_of_dfs, ~nest(.x))

all_the_categorical_variables <- 
  map_dfr(list_of_dfs, 
          ~gather(.x, variable, value, -cluster, -n))

# Compute proportions per category

flakes_clustered_cats_long <- 
all_the_categorical_variables %>% 
  mutate(value = str_squish(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  group_by(cluster, variable) %>%  
  mutate(prop = n / sum(n) * 100)  %>% 
  arrange(cluster, variable)
```

Here is the plot

```{r}
# sample sizes for each variable
flakes_clustered_cats_sample_sizes <- 
flakes_clustered_cats_long %>% 
  group_by(variable) %>% 
  summarise(tot_n = sum(n))

categorical_plots <- 
flakes_clustered_cats_long %>% 
  ungroup(cluster) %>% 
  nest(-variable) %>% 
  left_join(flakes_clustered_cats_sample_sizes) %>% 
  mutate(variable = glue::glue('{variable} (n = {tot_n})')) %>% 
  mutate(data = map(data, ~.x %>% mutate(value = as.character(value)))) %>% 
  mutate(plot = map2(data,
                     variable, ~.x %>%  
                            ggplot(aes(cluster, 
                                          prop, 
                                          fill = value)) + 
                       scale_fill_viridis_d() +
                            geom_col() +
                            xlab("") +
                            ylab("%") +
                       theme_minimal(base_size = 8) +
                           theme(legend.key.size = unit(0.4,"line"),
                                 legend.key.width = unit(0.4,"line"),
                                 legend.key.height = unit(0.4,"line"),
                                 legend.text = element_text(size=8), 
                                 legend.title = element_blank(),
                                 legend.margin=margin(t=-0.5, 
                                                      r=-0.5, 
                                                      b=-0.5, 
                                                      l=-0.1, 
                                                      unit="cm"),
                                 legend.box.margin=margin(-0.1,
                                                          -0.1,
                                                          -0.1,
                                                          -0.3,
                                                          unit="cm"),
                                 plot.margin = unit(x = c(0.5, 
                                                          0, 
                                                          0,
                                                          0.5), 
                                                    units = "cm")) +
                            ggtitle(str_glue('{.y}'))))

# arrange and draw the plots
categorical_plots_arranged <- 
cowplot::plot_grid(plotlist = categorical_plots$plot,
                   align = "hv",
                   axis = "tblr",
                   ncol = 2)

save_plot("analysis/report/Extended_Data_Figure_Y.png", categorical_plots_arranged,
          base_aspect_ratio = 1,
          base_height = 8,
          base_width = 18.3/2.5
)

```

A closer look at the distribution of retouch types by size classes

```{r eval=FALSE}
flakes_clustered_cats_long %>% 
  filter(variable == "Retouch type") %>% 
  ggplot(aes(value, prop)) +
  geom_col() +
  facet_wrap( ~ cluster , ncol = 1)
```

A closer look at the distribution of retouch types by size classes

```{r eval=FALSE}
flakes_clustered_cats_long %>% 
  filter(variable == "Retouch type") %>% 
  ggplot(aes(value, prop)) +
  geom_col() +
  facet_wrap( ~ cluster , ncol = 1)
```

Try this on the Lev flakes only:

```{r  eval=FALSE}
flakes_clustered %>% 
  select_if(function(col) is.numeric(col) | 
                                   all(col == .$type1))  %>% 
  filter(str_detect(type1, "LV|leva")) %>% 
    select(-box, -type1) %>% 
  gather(variable, value, -cluster) %>% 
  filter(!is.na(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  ggplot(aes(cluster,
             value)) +
  geom_boxplot() +
  facet_wrap( ~ variable, scales = "free_y") +
  theme_minimal()
```

```{r  eval=FALSE}
flakes_clustered_cats_long_leva <- 
flakes_clustered_cats %>% 
  filter(type1 %in% c("levallois", "LVF")) %>% 
  gather(variable, value, -cluster)  %>% 
  filter(!is.na(value), 
         variable != "comment",
         variable != "heat_damage",
         variable != "platform_preparation",
         variable != "type2c") %>% 
  mutate(value = str_squish(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  group_by(cluster, variable, value) %>% 
  tally() %>% 
  ungroup() %>% 
  group_by(cluster, variable) %>%  
  mutate(prop = n / sum(n) * 100)

categorical_plots_leva <- 
flakes_clustered_cats_long_leva %>% 
  ungroup(cluster) %>% 
  nest(-variable) %>% 
  mutate(plot = map(data , ~.x %>%  
                            ggplot(aes(cluster, 
                                          prop, 
                                          group=1, 
                                          fill = value)) + 
                            geom_col() +
                            xlab("") +
                           theme(legend.key.size = unit(0.4,"line"),
                                 legend.key.width = unit(0.2,"line"),
                                 legend.key.height = unit(0.4,"line"),
                                 legend.text = element_text(size=8))))

# get legend labels so we know what variable is plotted
# in each plot, not sure why I can't do this with map
for(i in 1:nrow(categorical_plots_leva)){
  categorical_plots_leva$plot[[i]] <- 
    categorical_plots_leva$plot[[i]] + guides(fill=guide_legend(title=unique(categorical_plots_leva$variable)[i]))
}

# arrange and draw the plots
cowplot::plot_grid(plotlist = categorical_plots_leva$plot,
                   align = "hv",
                   axis = "tblr")

```

### 2020-10-04 analysis/paper/paper.Rmd
tidy up and polish fig 1

---
title: "Supplementary Information for Diverse lithic production strategies in southwest China during Late Middle Pleistocene"
author:
  - Myfirstname Mylastname:
      email: fl@oneplace.edu
      institute: [UofO]
      correspondence: true
  - Yourfirstname Yourlastname:
      email: fl@another.edu
      institute: [UofA]
      correspondence: false
institute:
  - UofO: University of One Place
  - UofA: University of Another Place
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

<!-- The actual document text starts here: -->

Read in the data.

```{r get-data}
# Note the path that we need to use to access our data files when rendering this document
library(readxl)

file_name <- here::here("analysis/data/raw_data/artefacts-after Sam_Dec2018_quina_modified.xls")

flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
debris <- as.data.frame(read_excel(file_name, sheet = "chunk&debris"))
retouch <- as.data.frame(read_excel(file_name, sheet = "retouch"))
core_flake_scars_1 <- as.data.frame(read_excel(file_name, sheet = "core flake scars"))
edge_angle_1 <- as.data.frame(read_excel(file_name, sheet = "edge angle"))
retouched_pebble_chunk <- as.data.frame(read_excel(file_name, sheet = "Pebble&chunk"))

scar_dir <- as.data.frame(read_excel(file_name, sheet = 'scar direction'))

core_flake_scars <- core_flake_scars_1[core_flake_scars_1$number %in% cores$number, ]

edge_angle <- edge_angle_1[edge_angle_1$number %in% retouch$number, ]
```

Prepare and clean the data

```{r}
# prepare the data --------------------------------------------------

require(Ckmeans.1d.dp)
result <- Ckmeans.1d.dp(flakes$mass[!is.na(flakes$mass)])

n_clusters <- length(result$size)
n_clusters_30 <- length(result$size[result$size > 30])

# assign cluster ID to table
flakes_clustered <-
  flakes %>%
  filter(!is.na(mass)) %>%
  mutate(cluster = result$cluster) %>%
  right_join(flakes) %>%
  filter(cluster %in% 1:n_clusters_30)

# clean data to help with analysis
flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")
flakes <- flakes[!flakes$material %in% c('sandstone', 'quartz'),]
```

Results of descriptive statistics for Levallois and non-Levallois flakes. ‘PLF’ stands for preferential Levallois flake; ‘CF’ stands for complete flake.

```{r}
point_alpha <- 0.1
```

## Comparison of Levallois flakes vs non-Levallois flakes

Separate Levallois and Non-Levallois flakes for comparison

```{r}
library(tidyverse)
library(ggbeeswarm)
library(papaja)

# clean data to help with analysis
flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")
flakes <- flakes[!flakes$material %in% c('sandstone', 'quartz'),]

# how many levallois?
# unique(flakes$type1)

leva <- flakes %>% 
  filter(grepl("LVT|LVF|LVFB", type1)) %>%
  filter(!grepl("LVF_", type1))

non_leva <- flakes %>% 
  filter(!grepl("LVT|LVF|LVFB", type1))

flakes_L <-  
  mutate(flakes, L = ifelse(grepl("LVT|LVF|LVFB", type1), "Levallois", "Non-Levallois"))
```

Are the levallois pieces bigger by mass?

```{r}
# is it significant?
t_test_str_l_mass <- 
apa_print(t.test(data = flakes_L, mass ~ L))$statistic %>% 
  str_remove_all("\\$|\\{|\\}|\\\\|mathit")

# plot
compare_lev_non_lev_mass_plot <- 
ggplot(flakes_L, aes(x = L, 
                     y = mass)) +
  geom_boxplot() +
  geom_quasirandom(alpha = point_alpha) +
  scale_y_log10() + 
  theme_bw() + 
  xlab("") +
  ylab("Mass (g)") +
  ggtitle(paste0("Flake Mass: ", t_test_str_l_mass))
```

Are the levallois pieces bigger by max dimension?

```{r}
# is it significant?
t_test_str_l_maxdim <- 
apa_print(t.test(data = flakes_L, 
                 max_dimension ~ L))$statistic %>% 
  str_remove_all("\\$|\\{|\\}|\\\\|mathit")

# plot
compare_lev_non_lev_maxdim_plot <- 
ggplot(flakes_L, aes(x = L, 
                     y = max_dimension)) +
  geom_boxplot() +
  geom_quasirandom(alpha = point_alpha) +
  scale_y_log10() + 
  theme_bw() + 
  xlab("") +
  ylab("Maximum Dimension (mm)") +
  ggtitle(paste0("Flake Max. Dim.: ", t_test_str_l_maxdim))
```

Are the levallois pieces bigger by thickness?

```{r}
# is it significant?
t_test_str_l_thick <- 
apa_print(t.test(data = flakes_L,
                 `Thickness_50%max` ~ L))$statistic %>% 
  str_remove_all("\\$|\\{|\\}|\\\\|mathit")

# plot
compare_lev_non_lev_thick_plot <- 
ggplot(flakes_L, aes(x = L, 
                     y = `Thickness_50%max`)) +
  geom_boxplot() +
  geom_quasirandom(alpha = point_alpha) +
  scale_y_log10() + 
  theme_bw() + 
  xlab("") +
  ylab("Thickness (mm)") +
  ggtitle(paste0("Flake Thickness: ", t_test_str_l_thick))
```

## Comparison of Quina tools vs non-Quina tools

Do the Quina tools have different edge angles?

```{r}
library(reshape2)
edge_angle_2 <- edge_angle

edge_angle_2[, 2:(ncol(edge_angle_2)-1)] <- atan(edge_angle[, 2:(ncol(edge_angle_2)-1)]/2/3)/pi*180*2

df_angle <- melt(edge_angle_2[, 1:(ncol(edge_angle_2)-1)])

df_angle$variable <- substr(df_angle$variable, 1, 9)

colnames(df_angle)[2:3] <- c("section", "angle")

quina_id <- flakes$number[which(flakes$type3 == "quina")]

df_angle$quina <- ifelse(df_angle$number %in% quina_id, "Quina", "Non-Quina")

# siginificance test
t_test_str_q_angle <- 
df_angle %>% 
  drop_na("angle") %>% 
  t.test(data = ., angle ~ quina) %>% 
  apa_print %>% 
  .$statistic %>% 
  str_remove_all("\\$|\\{|\\}|\\\\|mathit")

compare_qui_non_qui_angle_plot <-
  ggplot(df_angle,
         aes(y = angle,
             x = quina)) +
  geom_boxplot() +
  geom_quasirandom(alpha = point_alpha) +
  theme_bw() +
  ylab("Edge angle (\u00B0)") +
  xlab("") +
  ggtitle(paste0("Flake Edge Angle: ", t_test_str_q_angle) )

```

Do the Quina tools have different thickness?

```{r}
df_thick <- melt(flakes[, c("number", "Thickness_25%max", "Thickness_50%max", "Thickness_75%max")])

df_thick$quina <- ifelse(df_thick$number %in% quina_id, "Quina", "Non-Quina")

df_thick$label <- substring(df_thick$variable, 11, 13)

df_thick_tt_test <- 
df_thick %>% 
  nest_by(label) %>% 
  mutate(t_test = list(
    t.test(data = data, value ~ quina) %>% 
      apa_print %>% 
      .$statistic %>% 
  str_remove_all("\\$|\\{|\\}|\\\\|mathit")
  ))

compare_qui_non_qui_thick_plot <-
  ggplot(df_thick,
         aes(y = value,
             x = quina)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.05) +
  geom_text(data = df_thick_tt_test,
            aes(label = t_test,
                x = 1.5, 
                y = 60)) +
  theme_bw() +
  facet_wrap(~label) +
  ylab("Thickness (mm)") + 
  xlab("")

```

Do the Quina tools have different GIUR?

```{r}
# plot GIUR on Quina  --------------------------------------------------

## calculate GIUR

# get tn
df_giur_tn <- retouch[, c(1,8,10,12,14,16,18,20,22)]

# divide tn by Tn
df_giur <- df_giur_tn[, -1] / retouch[, c(9,11,13,15,17,19,21,23)]

# compute GIUR as average of all zones
df_giur_values <- 
df_giur %>% 
  as_tibble %>% 
  rowwise %>% 
  mutate(GIUR = mean(c_across(section_1_t:section_8_t), 
                     na.rm = TRUE)) %>% 
  select(GIUR)

df_giur_values$number <- df_giur_tn$number

df_giur_clustered <- 
  left_join(df_giur_values, 
            flakes_clustered[, c("number", "cluster")]) %>% 
  drop_na() %>% 
  filter(GIUR <= 1)

df_giur_clustered$quina <- ifelse(df_giur_clustered$number %in% quina_id, "Quina", "Non-Quina")

df_giur_clustered_tt_test <- 
df_giur_clustered %>% 
    t.test(data = ., 
           GIUR ~ quina) %>% 
      apa_print %>% 
      .$statistic %>% 
  str_remove_all("\\$|\\{|\\}|\\\\|mathit")

# code for raincloud plot
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")

compare_qui_non_qui_giur_plot <-
  ggplot(df_giur_clustered) +
  aes(y = GIUR,
      x = quina) +
  geom_flat_violin(position = position_nudge(x = .25, 
                                             y = 0), 
                   alpha = .8,
                   adjust = 0.35)  +
  geom_quasirandom( 
             size = .75, 
             width = 0.25,
             alpha = point_alpha) +
  geom_boxplot(
        width = .25, 
               outlier.shape = NA, 
               alpha = 0.25
               ) +
  ggtitle(paste0("Flake GIUR: ", df_giur_clustered_tt_test)) +
  coord_flip() +
  theme_bw() +
  ylab("GIUR") +
  xlab("")

```

Plot combining multiple panels

```{r}
## plot figures together.
library(cowplot)
figure_leva_quina <-
  cowplot::plot_grid(plot_grid(compare_lev_non_lev_mass_plot, 
                               compare_lev_non_lev_maxdim_plot, 
                               compare_lev_non_lev_thick_plot ,
                               labels = "auto",
                               label_size = 12,
                               ncol = 3),
                     plot_grid( 
                               compare_qui_non_qui_thick_plot ,
                               labels = c("d"),
                               label_size = 12,
                               ncol = 1,
                               rel_widths = c(1,2)),
                     plot_grid(compare_qui_non_qui_angle_plot,
                               compare_qui_non_qui_giur_plot , 
                               labels= c("e", "f"),
                               ncol = 2,
                               #rel_widths = c(1.5, 1),
                               label_size = 12),
                     nrow = 3,
                     align = 'v',
                     axis = 'lr',
                     vjust = 1.5,
                     hjust = -0.5,
                     scale = 0.9) #

save_plot(here::here("analysis/figures/fig-1-Figures-for-Leva-and-quina.png"),
                     figure_leva_quina,
          base_aspect_ratio = 1,
          base_height = 10, #18,
          base_width =  14) #22)
```

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2020-10-04 analysis/paper/old-stuff/fig-1-compare-levallois-non-levallois.R

@@ -202,9 +202,9 @@ p_thick_quina <-
 
 ## calculate GIUR
 
-df_giur <- retouch[, c(1,8,10,12,14,16,18,20,22)]
+df_giur_tn <- retouch[, c(1,8,10,12,14,16,18,20,22)]
 
-df_giur[, -1] <- df_giur[, -1] / retouch[, c(9,11,13,15,17,19,21,23)]
+df_giur[, -1] <- df_giur_tn[, -1] / retouch[, c(9,11,13,15,17,19,21,23)]
 
 df_giur_clustered <- left_join(df_giur, flakes_clustered[, c("number", "cluster")])
 
@@ -215,7 +215,8 @@ df_giur_clustered$quina <- ifelse(df_giur_clustered$number %in% quina_id, "Quina
 df_giur_clustered_melt_2 <- melt(df_giur_clustered[,-1], id.vars = "quina")
 
 p_GIUR_quina <-
-  ggplot(na.omit(df_giur_clustered_melt_2[df_giur_clustered_melt_2$value <=1,]), aes(value)) +
+  ggplot(na.omit(df_giur_clustered_melt_2[df_giur_clustered_melt_2$value <=1,]),
+         aes(value)) +
   geom_histogram(binwidth = 0.1) +
   theme_bw() +
   theme(text = element_text(size = 20)) +

### 2020-10-04 analysis/paper/paper.Rmd
nice looking table 1

@@
 
-Results of descriptive statistics for Levallois and non-Levallois flakes. ‘PLF’ stands for preferential Levallois flake; ‘CF’ stands for complete flake.
-
-```{r}
-point_alpha <- 0.1
-```
-
-## Comparison of Levallois flakes vs non-Levallois flakes
-
 Separate Levallois and Non-Levallois flakes for comparison
@@
 flakes$retouched <- ifelse(grepl("ret", flakes$type1), "retouched", "unretouched")
-flakes <- flakes[!flakes$material %in% c('sandstone', 'quartz'),]
+flakes <-
+  flakes[!flakes$material %in% c('sandstone', 'quartz'), ] %>%
+  mutate(complet_flk = ifelse(
+    type1 %in% c(
+      "flake",
+      "blade",
+      "KBW",
+      "tablet",
+      "leva",
+      "leva?",
+      "DBD",
+      "crest"
+    ),
+    "complete flake",
+    ifelse(
+      type1 %in% c("flake brk", "blade brk", "LVFB", "proximal"),
+      "flake breaks",
+      "other"
+    )
+  ))
 
@@
 flakes_L <-  
-  mutate(flakes, L = ifelse(grepl("LVT|LVF|LVFB", type1), "Levallois", "Non-Levallois"))
+  mutate(flakes, L = ifelse(grepl("LVT|LVF|LVFB", type1), 
+                            "Levallois", "Non-Levallois"))  
 ```
+
+Results of descriptive statistics for Levallois and non-Levallois flakes. ‘PLF’ stands for preferential Levallois flake; ‘CF’ stands for complete flake.
+
+```{r}
+flakes_complete_non_Lev <- 
+  flakes_L %>% # only complete flakes for non-Levallois
+  filter(complet_flk == "complete flake")
+
+flakes_Lev  <- 
+  flakes_L %>% 
+  filter(L == "Levallois")
+
+# combine both here
+flakes_Lev_non_Lev <- 
+bind_rows(flakes_complete_non_Lev,
+          flakes_Lev) %>% 
+  as_tibble()
+
+# check the CVs
+CV <- function(the_vector){
+  mean_ <- mean(the_vector, na.rm = TRUE)
+  sd_ <- sd(the_vector, na.rm = TRUE)
+  cv <- (sd_/mean_) * 100
+  round(cv, 3)
+}
+
+# prepare mean and sd functions
+mean_t <- function(x) mean(x, na.rm = TRUE)
+sd_t <- function(x) sd(x, na.rm = TRUE)
+
+# for the table 'Results of descriptive statistics for Levallois and non-Levallois flakes. ‘PLF’ stands for preferential Levallois flake; ‘CF’ stands for complete flake.'
+compare_lev_non_lev_vars_tbl <- 
+flakes_Lev_non_Lev %>% 
+  as_tibble() %>% 
+  select(L, where(is.numeric)) %>% 
+  group_by(L) %>% 
+  summarise(across(where(is.numeric),
+            list(mean = mean_t,
+                 sd = sd_t,
+                 cv = CV),
+            .names = "{.fn}.{.col}")) %>%
+  pivot_longer(-L) %>%
+  separate(col = name,
+           into = c("stat", "var"),
+           sep = "\\.") %>% 
+  mutate(L = ifelse(L == "Levallois",
+                    "PLF",
+                    "CF")) %>% 
+  unite("lev_stat", c(stat, L), 
+        sep = "_") %>% 
+  pivot_wider(names_from = lev_stat ) %>% 
+  mutate(difference_in_cv = cv_PLF - `cv_CF`) %>% 
+  mutate(across(where(is.numeric), round, 2))
+
+library(gt) 
+tab <- 
+  compare_lev_non_lev_vars_tbl %>%
+  gt(rowname_col = "var") %>% 
+  tab_spanner(
+    label = "Mean (mm)",
+    columns = vars(mean_PLF, 
+                   mean_CF)) %>% 
+  tab_spanner(
+    label = "SD",
+    columns = vars(sd_PLF, 
+                   sd_CF)  
+  ) %>% 
+  tab_spanner(
+    label = "CV (%)",
+    columns = vars(cv_PLF, 
+                   cv_CF,
+                   difference_in_cv) 
+  )
+
+```
+
+Are CV values lower for Levallois flakes than for non-Levallois flakes?
+
+```{r}
+compare_lev_non_lev_vars_tbl  %>% 
+  select(contains("cv_")) %>% 
+  stack() %>% 
+  wilcox.test(values ~ ind, 
+              data = .) %>% 
+  apa_print() %>% 
+  .$statistic %>% 
+  str_remove_all("\\$|\\{|\\}|\\\\|mathit")
+
+```
+
+## Comparison of Levallois flakes vs non-Levallois flakes
 
@@
 ```{r}
+# plotting setting
+point_alpha <- 0.1
+
 # is it significant?

### 2020-10-04 analysis/paper/old-stuff/script for paper II_2.R

@@ -441,7 +441,7 @@ flake_summary <- rbind(flake_summary,
 
 print("table for flake attributes")
 
-kable(round(flake_summary, digits = 1))
+knitr::kable(round(flake_summary, digits = 1))
 
 ## how many flake type1
 

### 2020-10-06 analysis/paper/paper.Rmd
check SI

@@
 title: "Supplementary Information for Diverse lithic production strategies in southwest China during Late Middle Pleistocene"
-author:
-  - Myfirstname Mylastname:
-      email: fl@oneplace.edu
-      institute: [UofO]
-      correspondence: true
-  - Yourfirstname Yourlastname:
-      email: fl@another.edu
-      institute: [UofA]
-      correspondence: false
-institute:
-  - UofO: University of One Place
-  - UofA: University of Another Place
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: 
-    bookdown::word_document2:
-      fig_caption: yes
-      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
-      pandoc_args:
-      - --lua-filter=../templates/scholarly-metadata.lua
-      - --lua-filter=../templates/author-info-blocks.lua
-      - --lua-filter=../templates/pagebreak.lua
-bibliography: references.bib
-csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
-abstract: |
-  Text of abstract
-keywords: |
-  keyword 1; keyword 2; keyword 3
-highlights: |
-  These are the highlights. 
 ---
@@
   message = FALSE,
-  echo = FALSE,
+  echo = TRUE,
   comment = "#>",
@@
 library(readxl)
+library(tidyverse)
 
@@
 
+tab
+
 ```
@@
   ggtitle(paste0("Flake Mass: ", t_test_str_l_mass))
+
+compare_lev_non_lev_mass_plot
 ```
@@
   ggtitle(paste0("Flake Max. Dim.: ", t_test_str_l_maxdim))
+
+compare_lev_non_lev_maxdim_plot
 ```
@@
   ggtitle(paste0("Flake Thickness: ", t_test_str_l_thick))
+
+compare_lev_non_lev_thick_plot
 ```
@@
 
+compare_qui_non_qui_angle_plot
+
 ```
@@
   xlab("")
+
+compare_qui_non_qui_thick_plot
 
@@
 
+compare_qui_non_qui_giur_plot
+
 ```
@@
           base_width =  14) #22)
+
+figure_leva_quina
 ```
-
-<!-- The following line inserts a page break  -->
-
-\newpage
-
-# References
-
-<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
-
-<div id="refs"></div>
 

### 2020-10-06 analysis/paper/supplementary-information.R

@@ -1,0 +1,1351 @@
+
+#install required packages
+library(dplyr)
+library(knitr)
+library(ggplot2)
+library(reshape2)
+library(readxl)
+library(fitdistrplus)
+library(tidyr)
+library(classInt)
+library(cowplot)
+library(gridExtra)
+
+# read in the data
+
+
+file_name <- here::here("analysis/data/raw_data/artefacts-after Sam_May2020_quina_modified.xls")
+
+flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
+cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
+debris <- as.data.frame(read_excel(file_name, sheet = "chunk&debris"))
+retouch <- as.data.frame(read_excel(file_name, sheet = "retouch"))
+core_flake_scars_1 <- as.data.frame(read_excel(file_name, sheet = "core flake scars"))
+edge_angle_1 <- as.data.frame(read_excel(file_name, sheet = "edge angle"))
+retouched_pebble_chunk <- as.data.frame(read_excel(file_name, sheet = "Pebble&chunk"))
+
+scar_dir <- as.data.frame(read_excel(file_name, sheet = 'scar direction'))
+
+core_flake_scars <- core_flake_scars_1[core_flake_scars_1$number %in% cores$number, ]
+
+edge_angle <- edge_angle_1[edge_angle_1$number %in% retouch$number, ]
+
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+# basic details of the dataset
+
+artefact_ids <- list(flake_ids = flakes$number,
+                     core_ids = cores$number,
+                     debris_ids = debris$number)
+
+retouch_ids <- retouch$number
+
+# how many unique artefacts?
+
+totals_of_each_type <- lapply(artefact_ids, function(i) length(i))
+
+count_unique_artefacts <- sum(unlist(totals_of_each_type))
+
+# debris pieces that are retouch pieces
+
+debris_with_retouch <- intersect(retouch_ids, artefact_ids$debris_ids)
+
+proportion_debris_retouched <- length(debris_with_retouch) / count_unique_artefacts
+
+# proportion of each type
+
+type_table <- data.frame(type = names(as.data.frame(totals_of_each_type)),
+                         count = as.numeric(as.data.frame(totals_of_each_type)),
[…]

### 2021-06-16 analysis/paper/supplementary-information.R
working through the SI

@@ -1,6 +1,7 @@
 
 #install required packages
 library(dplyr)
+library(readr)
 library(knitr)
 library(ggplot2)
 library(reshape2)
@@ -12,7 +13,6 @@ library(cowplot)
 library(gridExtra)
 
 # read in the data
-
 
 file_name <- here::here("analysis/data/raw_data/artefacts-after Sam_May2020_quina_modified.xls")
 
@@ -66,10 +66,9 @@ type_table <- rbind(type_table, c("total", count_unique_artefacts, round(sum(typ
 
 # print pretty table
 
-print("table of lithic types")
+# print("table of lithic types")
 
 kable(type_table)
-
 
 ### raw materials
 
@@ -94,9 +93,7 @@ res <- flakes %>%
   replace(., is.na(.), 0) %>%
   rename(category = complet_flk)
 
-
 raw_material_table <- bind_rows(raw_material_table, res[1:2,])
-
 
 res <- cores %>% filter(!is.na(material)) %>%
   dplyr :: count(material) %>%
@@ -105,14 +102,12 @@ res <- cores %>% filter(!is.na(material)) %>%
 
 raw_material_table <- bind_rows(raw_material_table, res)
 
-
 res <- debris %>% filter(!is.na(material)) %>%
   dplyr :: count(material) %>%
   spread(material, n) %>%
   mutate(category = "debris")
 
 raw_material_table <- bind_rows(raw_material_table, res)
-
 
 res <- flakes %>%
   filter(type1 %in% c("ret", "ret brk", "BUR", "LVT", "ret leva?", "ret pro-lepoin")) %>%
@@ -122,7 +117,6 @@ res <- flakes %>%
   mutate(category = "retouched flakes and breaks")
 
 raw_material_table <- bind_rows(raw_material_table, res)
-
 
 res <- retouched_pebble_chunk %>%
 dplyr :: count(material) %>%
@@ -175,27 +169,58 @@ res <- flakes %>%
 
 raw_material_table <- bind_rows(raw_material_table, res)
 
-
 raw_material_table[is.na(raw_material_table)] <- 0 ##convert na to 0
 
 raw_material_table$total <- rowSums(raw_material_table[,-1])
-
-write.csv(raw_material_table, "raw_material_table.csv")
-
 
 x <- raw_material_table
 
 prop_raw_material <- cbind(category = x[, 1], round(x[, -1]/rowSums(x[, 2:(ncol(x)-1)])*100, digits = 1))
 
[…]

### 2021-06-25 analysis/paper/paper.Rmd
working through the SI
updates and checking of manuscript and SI

@@
 
-file_name <- here::here("analysis/data/raw_data/artefacts-after Sam_Dec2018_quina_modified.xls")
+file_name <-
+  here::here("analysis/data/raw_data/artefacts-after Sam_Dec2018_quina_modified-2019.xls")
 
@@
 
-Results of descriptive statistics for Levallois and non-Levallois flakes. ‘PLF’ stands for preferential Levallois flake; ‘CF’ stands for complete flake.
+Results of descriptive statistics for Levallois and non-Levallois flakes. ‘PLF’ stands for preferential Levallois flake; ‘CF’ stands for complete flake. This is table 1 in the main text
 
@@
 
-Are CV values lower for Levallois flakes than for non-Levallois flakes?
+Are CV values lower for Levallois flakes than for non-Levallois flakes? This result is presented in line in the text in the main text.
 
@@
 ## calculate GIUR
+# looking at the GIUR data... there are some duplicate artefact numbers, which is a nuisance
 
-# get tn
-df_giur_tn <- retouch[, c(1,8,10,12,14,16,18,20,22)]
+# make the number ID unique
+retouch$number <- make.unique(as.character(retouch$number), sep = "_")
 
-# divide tn by Tn
-df_giur <- df_giur_tn[, -1] / retouch[, c(9,11,13,15,17,19,21,23)]
+retouch_giur_per_section <-
+  retouch %>%
+  dplyr::select(grep("^number$|_t|_T", names(retouch))) %>%
+  gather(variable, value, -number) %>%
+  separate(variable, c("section", "tee"), sep = 10) %>%
+  pivot_wider(names_from =  tee, values_from = value) %>%
+  mutate(t_T = t / T)  %>%
+  filter(t_T <= 1)
 
-# compute GIUR as average of all zones
-df_giur_values <- 
-df_giur %>% 
-  as_tibble %>% 
-  rowwise %>% 
-  mutate(GIUR = mean(c_across(section_1_t:section_8_t), 
-                     na.rm = TRUE)) %>% 
-  select(GIUR)
-
-df_giur_values$number <- df_giur_tn$number
-
-df_giur_clustered <- 
-  left_join(df_giur_values, 
-            flakes_clustered[, c("number", "cluster")]) %>% 
-  drop_na() %>% 
-  filter(GIUR <= 1)
-
-df_giur_clustered$quina <- ifelse(df_giur_clustered$number %in% quina_id, "Quina", "Non-Quina")
+retouch_giur_per_artefact <-
+  retouch_giur_per_section %>%
+  group_by(number) %>%
+  summarise(mean_giur = mean(t_T, na.rm = TRUE)) %>% 
+  mutate(quina = ifelse(number %in% quina_id, "Quina", "Non-Quina"))
 
 df_giur_clustered_tt_test <- 
-df_giur_clustered %>% 
+retouch_giur_per_artefact %>% 
     t.test(data = ., 
-           GIUR ~ quina) %>% 
+           mean_giur ~ quina) %>% 
       apa_print %>% 
@@
 compare_qui_non_qui_giur_plot <-
-  ggplot(df_giur_clustered) +
-  aes(y = GIUR,
+  ggplot(retouch_giur_per_artefact) +
+  aes(y = mean_giur,
       x = quina) +
@@
   geom_quasirandom( 
-             size = .75, 
+             size = 1.5, 
              width = 0.25,
@@
 
-Plot combining multiple panels
+Plot combining multiple panels, this is Figure 1 is the main text
 

### 2021-06-25 analysis/paper/supplementary-information.R

@@ -14,7 +14,8 @@ library(gridExtra)
 
 # read in the data
 
-file_name <- here::here("analysis/data/raw_data/artefacts-after Sam_May2020_quina_modified.xls")
+file_name <-
+  here::here("analysis/data/raw_data/artefacts-after Sam_Dec2018_quina_modified-2019.xls")
 
 flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
 cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
@@ -322,15 +323,24 @@ print("summary for core platform preparation")
 # what are these platform types?
 unique(cores$platform_preparation)
 
-# cores <-
-#   cores %>%
-#   mutate(platform_preparation = case_when(
-#
-#
-#   ))
+cores <-
+  cores %>%
+  mutate(platform_preparation = case_when(
+    platform_preparation == "over" ~ 'overhang\nremoval',
+    platform_preparation == "plane" ~ 'plain',
+    platform_preparation == "faceted" ~ 'facetted',
+    platform_preparation == "cortex" ~ 'cortical',
+    platform_preparation == "dehid" ~ 'dihedral'
+  )) %>%
+  drop_na(platform_preparation)
 
+# "over"    "plain"   "plane"   "faceted"  "n"    "cortex"  "dehid".
+# 'overhang removal', 'plain' (same as plane?), 'facetted', 'cortical', 'dihedral'
 
-p_core_plat_prep <- f(cores, "platform_preparation") + xlab("Platform preparation") + ylab("Count")
+p_core_plat_prep <- f(cores, "platform_preparation") +
+  xlab("Platform preparation") +
+  ylab("Count") +
+  coord_flip()
 
 core_plat_type <-  table(na.omit(cores$platform_preparation))
 
@@ -421,7 +431,7 @@ figure_cores <-
                      hjust = -0.5,
                      scale = 0.95) #
 
-save_plot("Figures for cores.png", figure_cores,
+save_plot("analysis/figures/S2 figures for cores.png", figure_cores,
           base_aspect_ratio = 1,
           base_height = 13,
           base_width = 21)
@@ -641,10 +651,13 @@ p_flake_scar_number <- ggplot(flakes, aes(`scar_number`)) +
         axis.title.x = element_text(size = 25),
         axis.title.y = element_text(size = 25))
 
[…]

### 2021-06-25 README.Rmd
rearrange files slightly

@@
 # Please put your title here to include it in the file below.
-Title <- "Title of your paper goes here"
+Title <- "Diverse lithic production strategies in southwest China during Late Middle Pleistocene "
 ```
@@
 
-> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>
+> Yue Hu, Ben Marwick, (nd). _`r Title`_.  <https://doi.org/xxx/xxx>
 
@@
 
-> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+> Yue Hu, Ben Marwick, (nd). _`r Title`_. , Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
 
@@
 
-> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+> Yue Hu, Ben Marwick, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
 
@@
     illustrations
-  - [:file\_folder:
-    supplementary-materials](/analysis/supplementary-materials):
-    Supplementary materials including notes and other documents
-    prepared and collected during the analysis.
 

### 2021-06-25 analysis/paper/old-stuff/old files/for_paper_ii.tex
rearrange files slightly

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{}
    \pretitle{\vspace{\droptitle}}
  \posttitle{}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\begin{document}

Some normal text.

\begin{verbatim}
## [1] "table of lithic types"
\end{verbatim}

\begin{longtable}[]{@{}lll@{}}
\toprule
type & count & proportion\tabularnewline
\midrule
\endhead
flake\_ids & 1199 & 0.541\tabularnewline
core\_ids & 249 & 0.112\tabularnewline
debris\_ids & 769 & 0.347\tabularnewline
total & 2217 & 1\tabularnewline
\bottomrule
\end{longtable}

\begin{verbatim}
## [1] "tables for raw materials of the assemblage"
\end{verbatim}

\begin{longtable}[]{@{}lrrrrrr@{}}
\toprule
category & chert & limestone & basalt & quartz & sandstone &
total\tabularnewline
\midrule
\endhead
flake & 932 & 249 & 11 & 2 & 5 & 1199\tabularnewline
core & 208 & 38 & 2 & 0 & 0 & 248\tabularnewline
debris & 569 & 190 & 8 & 1 & 1 & 769\tabularnewline
retouched pieces & 767 & 201 & 11 & 1 & 4 & 984\tabularnewline
backed knife & 6 & 2 & 0 & 0 & 0 & 8\tabularnewline
bec & 6 & 1 & 0 & 0 & 0 & 7\tabularnewline
borer & 52 & 17 & 0 & 0 & 0 & 69\tabularnewline
BUR & 4 & 1 & 0 & 0 & 0 & 5\tabularnewline
chopper & 0 & 0 & 1 & 0 & 0 & 1\tabularnewline
cleaver & 1 & 0 & 0 & 0 & 0 & 1\tabularnewline
Denticular & 54 & 21 & 1 & 0 & 1 & 77\tabularnewline
Endscraper & 32 & 6 & 0 & 0 & 0 & 38\tabularnewline
Point & 24 & 5 & 0 & 0 & 0 & 29\tabularnewline
scraper & 486 & 130 & 7 & 1 & 3 & 627\tabularnewline
nature backed & 4 & 2 & 0 & 0 & 0 & 6\tabularnewline
notch & 65 & 12 & 2 & 0 & 0 & 79\tabularnewline
tanged point & 9 & 1 & 0 & 0 & 0 & 10\tabularnewline
\bottomrule
\end{longtable}

\begin{verbatim}
## [1] "proportion tables for raw materials of the assemblage"
\end{verbatim}

\begin{longtable}[]{@{}lrrrrrr@{}}
\toprule
category & chert & limestone & basalt & quartz & sandstone &
total\tabularnewline
\midrule
\endhead
flake & 77.7 & 20.8 & 0.9 & 0.2 & 0.4 & 100\tabularnewline
core & 83.9 & 15.3 & 0.8 & 0.0 & 0.0 & 100\tabularnewline
debris & 74.0 & 24.7 & 1.0 & 0.1 & 0.1 & 100\tabularnewline
retouched pieces & 77.9 & 20.4 & 1.1 & 0.1 & 0.4 & 100\tabularnewline
backed knife & 75.0 & 25.0 & 0.0 & 0.0 & 0.0 & 100\tabularnewline
bec & 85.7 & 14.3 & 0.0 & 0.0 & 0.0 & 100\tabularnewline
borer & 75.4 & 24.6 & 0.0 & 0.0 & 0.0 & 100\tabularnewline
BUR & 80.0 & 20.0 & 0.0 & 0.0 & 0.0 & 100\tabularnewline
chopper & 0.0 & 0.0 & 100.0 & 0.0 & 0.0 & 100\tabularnewline
cleaver & 100.0 & 0.0 & 0.0 & 0.0 & 0.0 & 100\tabularnewline
Denticular & 70.1 & 27.3 & 1.3 & 0.0 & 1.3 & 100\tabularnewline
Endscraper & 84.2 & 15.8 & 0.0 & 0.0 & 0.0 & 100\tabularnewline
Point & 82.8 & 17.2 & 0.0 & 0.0 & 0.0 & 100\tabularnewline
scraper & 77.5 & 20.7 & 1.1 & 0.2 & 0.5 & 100\tabularnewline
nature backed & 66.7 & 33.3 & 0.0 & 0.0 & 0.0 & 100\tabularnewline
notch & 82.3 & 15.2 & 2.5 & 0.0 & 0.0 & 100\tabularnewline
tanged point & 90.0 & 10.0 & 0.0 & 0.0 & 0.0 & 100\tabularnewline
\bottomrule
\end{longtable}

\begin{verbatim}
## [1] "summary of core attributes"
\end{verbatim}

\begin{longtable}[]{@{}lrrrrrrrrrrrr@{}}
\toprule
& length & max\_dimension & medial\_width & distal\_width & thickness &
distal\_thickness & mass & platform & platform\_width &
platform\_thickness & scar\_number & cortex\_percentage\tabularnewline
\midrule
\endhead
mean & 47.2 & 75.1 & 58.0 & 49.2 & 51.8 & 38.7 & 198.9 & 1.5 & 50.9 &
43.7 & 2.9 & 14.5\tabularnewline
sd & 20.2 & 21.6 & 21.2 & 21.5 & 29.6 & 17.1 & 166.8 & 0.8 & 20.0 & 19.8
& 2.0 & 19.4\tabularnewline
CV & 0.4 & 0.3 & 0.4 & 0.4 & 0.6 & 0.4 & 0.8 & 0.5 & 0.4 & 0.5 & 0.7 &
1.3\tabularnewline
25\% & 33.0 & 63.0 & 43.0 & 35.2 & 38.2 & 27.2 & 100.2 & 1.0 & 37.0 &
30.2 & 1.2 & 0.0\tabularnewline
50\% & 43.5 & 72.0 & 55.0 & 45.0 & 47.0 & 37.0 & 149.5 & 1.0 & 50.0 &
40.0 & 2.0 & 7.5\tabularnewline
75\% & 57.8 & 83.8 & 68.0 & 60.6 & 60.8 & 48.0 & 243.0 & 2.0 & 61.8 &
51.9 & 4.0 & 20.0\tabularnewline
\bottomrule
\end{longtable}

\begin{verbatim}
## [1] "cortex proportion of cores"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-2-1.pdf}

We will try univariate clustering for flake mass:

\begin{verbatim}
## [1] "flake grouping result according to mass"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-3-1.pdf}

\begin{verbatim}
##  [1] "crest"          "DBD"            "flake"          "KBW"           
##  [5] "leva"           "LVF"            "LVF_?"          "ret"           
##  [9] "ret brk"        "ret leva?"      "ret pro-lepoin" "tablet"        
## [13] "blade"          "leva?"          "blade brk?"     "BUR"           
## [17] "FACETED"        "flake brk"      "LVFB"           "LVT"           
## [21] "proximal"       NA
\end{verbatim}

\begin{verbatim}
## [1] "compare the mass of Leva and non-leva flakes"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-4-1.pdf}

\begin{verbatim}
## [1] "t-test to see if it is statistically significant?"
\end{verbatim}

\begin{verbatim}
## 
##  Welch Two Sample t-test
## 
## data:  mass by L
## t = -1.9714, df = 64.674, p-value = 0.05296
## alternative hypothesis: true difference in means is not equal to 0
## 95 percent confidence interval:
##  -28.6440377   0.1870471
## sample estimates:
## mean in group L mean in group N 
##        39.28571        53.51421
\end{verbatim}

\begin{verbatim}
## [1] "variation of thickness for all flakes (red = 25%max_dim,green = 50%max,blue=75%max"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-4-2.pdf}

\begin{verbatim}
## [1] "compare thickness of leva and non-leva (red = 25%max_dim, \n      green = 50%max, blue=75%max"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-4-3.pdf}

\begin{verbatim}
## [1] "summary of cv for thickness and width for leva and non-leva flakes"
\end{verbatim}

\begin{verbatim}
## # A tibble: 2 x 8
##   L     cv_25_thick cv_50_thick cv_75_thick cv_25_width cv_50_width
##   <chr>       <dbl>       <dbl>       <dbl>       <dbl>       <dbl>
## 1 L            40.2        40.0        42.7        38.3        36.0
## 2 N            48.8        44.7        46.0        79.3        40.6
## # ... with 2 more variables: cv_75_width <dbl>, n <int>
\end{verbatim}

The following are results for cores

\begin{verbatim}
## [1] "summary for core geometry"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-5-1.pdf}

\begin{verbatim}
## 
##   circle   column    conic    cubic  end,scp iregular   wedged 
##     0.61     6.71     9.76     0.61     0.61    80.49     1.22
\end{verbatim}

\begin{verbatim}
## [1] "summary for core platform"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-5-2.pdf}

\begin{verbatim}
## 
##     1     2     3     4     5    36 
## 60.49 27.16  6.17  4.94  0.62  0.62
\end{verbatim}

\begin{verbatim}
## [1] "summary for core platform preparation"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-5-3.pdf}

\begin{verbatim}
## 
##  cortex   dehid faceted       n    over   plain   plane 
##    3.38    0.68   18.24    6.76   15.54   54.05    1.35
\end{verbatim}

\begin{verbatim}
## [1] "summary for core rotations"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-5-4.pdf}

\begin{verbatim}
## 
##     0     1     2     3     4     5 
##  8.05 59.73 22.15  6.04  2.68  1.34
\end{verbatim}

\begin{verbatim}
## [1] "summary for core reduction type"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-5-5.pdf}

\begin{verbatim}
## 
## <U+7A84><U+9762><U+77F3><U+6838>           BIFACIAL      BIFACIAL CORE 
##               0.40               1.21               2.42 
##         blade core             blade?               core 
##               4.84               1.21               0.40 
##               disc             double               HMSF 
##               3.63              12.50               2.82 
##           KBW CORE             KBW/TF               leva 
##               0.81               3.23               0.40 
##                LVC            LVC/DIS   LVC?HMSF? DOUBLE 
##               4.44               0.40               0.40 
## LVMC(LV MIDDLE)/TF              multp         MULTP CORE 
##               0.40               5.65               0.40 
##             pebble          pre leva?             single 
##               0.40               0.40              31.05 
##        SINGLE CORE           spheroid                 TF 
##               0.81               0.40              21.37
\end{verbatim}

\begin{verbatim}
## [1] "summary of core scar number"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-5-6.pdf}

\begin{verbatim}
## 
##     1     2     3     4     5     6     7     8     9    10    14 
## 24.84 24.22 16.77 16.15  9.32  4.97  1.24  0.62  0.62  0.62  0.62
\end{verbatim}

The following results are for flakes

\begin{verbatim}
## [1] "table for flake attributes"
\end{verbatim}

\begin{longtable}[]{@{}lrrrrrrrrrrrrrrr@{}}
\toprule
& length & max\_dimension & oriented\_width & Width\_25\%max &
Width\_50\%max & Width\_75\%max & oriented\_thickness &
Thickness\_25\%max & Thickness\_50\%max & Thickness\_75\%max & mass &
platform\_width & platform\_thickness & scar\_number &
cortex\_percentage\tabularnewline
\midrule
\endhead
mean & 49.4 & 62.5 & 50.3 & 38.3 & 41.9 & 35.2 & 17.6 & 16.1 & 16.6 &
13.7 & 68.2 & 32.8 & 13.7 & 2.9 & 9.4\tabularnewline
sd & 19.2 & 22.5 & 19.2 & 38.6 & 15.3 & 14.8 & 8.1 & 7.6 & 7.8 & 6.8 &
81.7 & 16.8 & 7.8 & 1.6 & 15.1\tabularnewline
cv & 0.4 & 0.4 & 0.4 & 1.0 & 0.4 & 0.4 & 0.5 & 0.5 & 0.5 & 0.5 & 1.2 &
0.5 & 0.6 & 0.6 & 1.6\tabularnewline
25\% & 35.0 & 48.2 & 36.0 & 26.0 & 30.1 & 25.0 & 11.8 & 10.9 & 11.0 &
9.0 & 18.8 & 19.9 & 8.0 & 2.0 & 0.0\tabularnewline
50\% & 48.0 & 60.0 & 49.0 & 35.0 & 41.0 & 34.0 & 16.0 & 15.0 & 15.0 &
12.6 & 45.5 & 30.9 & 12.0 & 3.0 & 0.0\tabularnewline
75\% & 60.8 & 76.0 & 61.7 & 44.9 & 51.1 & 43.0 & 23.0 & 20.9 & 21.0 &
18.0 & 90.0 & 43.0 & 18.3 & 4.0 & 10.0\tabularnewline
\bottomrule
\end{longtable}

\begin{verbatim}
## [1] "summary for flake type1"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-1.pdf}

\begin{verbatim}
## 
##          blade     blade brk?            BUR          crest            DBD 
##              2              1              1              1             26 
##        FACETED          flake      flake brk            KBW           leva 
##              1            123              2             20             12 
##          leva?            LVF          LVF_?           LVFB            LVT 
##              1             30             21              3              3 
##       proximal            ret        ret brk      ret leva? ret pro-lepoin 
##              1            202            720              5              1 
##         tablet 
##              3
\end{verbatim}

\begin{verbatim}
## 
##          blade     blade brk?            BUR          crest            DBD 
##            0.2            0.1            0.1            0.1            2.2 
##        FACETED          flake      flake brk            KBW           leva 
##            0.1           10.4            0.2            1.7            1.0 
##          leva?            LVF          LVF_?           LVFB            LVT 
##            0.1            2.5            1.8            0.3            0.3 
##       proximal            ret        ret brk      ret leva? ret pro-lepoin 
##            0.1           17.1           61.1            0.4            0.1 
##         tablet 
##            0.3
\end{verbatim}

\begin{verbatim}
## [1] "summary for flake type2"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-2.pdf}

\begin{verbatim}
## 
##       backed knife                bec            bec,scp 
##                  8                  4                  3 
##              blade              borer         borer,dent 
##                  1                 13                  1 
##     borer,dent,scp      borer,end,scp        borer,notch 
##                  2                  1                  2 
##    borer,notch,scp          borer,scp      borer,scp,end 
##                  1                 47                  1 
##         borer,strt                BUR         BUR? CORE? 
##                  1                  3                  1 
##              burin              BURIN            cleaver 
##                  1                  2                  1 
## convergent sidescp                 cp              crest 
##                  1                  1                  1 
##               dent     dent,borer,scp           dent,end 
##                 51                  1                  1 
##         dent,notch           dent,scp           dent/NOT 
##                  2                 20                  1 
##         dent/POINT                end          end,borer 
##                  1                 14                  1 
##      end,borer,scp            end,scp             endscp 
##                  1                 17                  6 
##            kombewa               leva         leva point 
##                  1                  2                  1 
##        leva point?              leva?               moon 
##                  2                  1                  7 
##            NATBACK      nature backed             notcch 
##                  7                  1                  1 
##              notch        notch,borer          notch,end 
##                 61                  2                  2 
##          notch,scp         notch/DENT              point 
##                 21                  1                 22 
##              POINT        point break       point,broken 
##                  1                  1                  1 
##             point?          PSD POINT                scp 
##                  5                  1                606 
##            sidescp        sidescp-cvx             tablet 
##                 20                  1                  1 
##       tablet break               TANG             TANGED 
##                  1                  1                  1 
##       TANGED POINT  TANGED POINT /NOT      tanged point? 
##                  3                  1                  3 
##      TANGED POINT?      THINING POINT         TRANVS SCP 
##                  1                  1                  1
\end{verbatim}

\begin{verbatim}
## 
##       backed knife                bec            bec,scp 
##                0.8                0.4                0.3 
##              blade              borer         borer,dent 
##                0.1                1.3                0.1 
##     borer,dent,scp      borer,end,scp        borer,notch 
##                0.2                0.1                0.2 
##    borer,notch,scp          borer,scp      borer,scp,end 
##                0.1                4.7                0.1 
##         borer,strt                BUR         BUR? CORE? 
##                0.1                0.3                0.1 
##              burin              BURIN            cleaver 
##                0.1                0.2                0.1 
## convergent sidescp                 cp              crest 
##                0.1                0.1                0.1 
##               dent     dent,borer,scp           dent,end 
##                5.1                0.1                0.1 
##         dent,notch           dent,scp           dent/NOT 
##                0.2                2.0                0.1 
##         dent/POINT                end          end,borer 
##                0.1                1.4                0.1 
##      end,borer,scp            end,scp             endscp 
##                0.1                1.7                0.6 
##            kombewa               leva         leva point 
##                0.1                0.2                0.1 
##        leva point?              leva?               moon 
##                0.2                0.1                0.7 
##            NATBACK      nature backed             notcch 
##                0.7                0.1                0.1 
##              notch        notch,borer          notch,end 
##                6.1                0.2                0.2 
##          notch,scp         notch/DENT              point 
##                2.1                0.1                2.2 
##              POINT        point break       point,broken 
##                0.1                0.1                0.1 
##             point?          PSD POINT                scp 
##                0.5                0.1               61.0 
##            sidescp        sidescp-cvx             tablet 
##                2.0                0.1                0.1 
##       tablet break               TANG             TANGED 
##                0.1                0.1                0.1 
##       TANGED POINT  TANGED POINT /NOT      tanged point? 
##                0.3                0.1                0.3 
##      TANGED POINT?      THINING POINT         TRANVS SCP 
##                0.1                0.1                0.1
\end{verbatim}

\begin{verbatim}
## [1] "summary for flake retouch"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-3.pdf}

\begin{verbatim}
## 
##   retouched unretouched 
##         928         258
\end{verbatim}

\begin{verbatim}
## 
##   retouched unretouched 
##        78.2        21.8
\end{verbatim}

\begin{verbatim}
## [1] "summary for flake material"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-4.pdf}

\begin{verbatim}
## [1] "summary for flake platform shape"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-5.pdf}

\begin{verbatim}
## 
##   fusiform   gullwing   iregular quadrangle  rectangle  trapezoid 
##         46         31          1         87          1          4 
##   triangle 
##        136
\end{verbatim}

\begin{verbatim}
## 
##   fusiform   gullwing   iregular quadrangle  rectangle  trapezoid 
##      15.03      10.13       0.33      28.43       0.33       1.31 
##   triangle 
##      44.44
\end{verbatim}

\begin{verbatim}
## [1] "summary for flake platform"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-6.pdf}

\begin{verbatim}
## [1] "number of flakes with distinctive platform = 396"
\end{verbatim}

\begin{verbatim}
## 
##              cotex          dihederal            faceted 
##                 36                 45                 42 
## faceted(LEVA LIKE)           FACETED?              focus 
##                  1                  3                 20 
##               miss              overh              plain 
##                 35                  1                212 
##          RETOUCHED 
##                  1
\end{verbatim}

\begin{verbatim}
## 
##              cotex          dihederal            faceted 
##                9.1               11.4               10.6 
## faceted(LEVA LIKE)           FACETED?              focus 
##                0.3                0.8                5.1 
##               miss              overh              plain 
##                8.8                0.3               53.5 
##          RETOUCHED 
##                0.3
\end{verbatim}

\begin{verbatim}
## [1] "mass of flakes"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-7.pdf}
\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-8.pdf}

\begin{verbatim}
## [1] "length vs width at 50% max dim"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-9.pdf}

\begin{verbatim}
## [1] "platform shape vs thickness at 50% max dim"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-10.pdf}

\begin{verbatim}
## [1] "mass for retouched and non-retouched flakes"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-11.pdf}

\begin{verbatim}
## [1] "mass for flakes of different raw materials"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-12.pdf}

\begin{verbatim}
## [1] "max dim for all flakes"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-13.pdf}

\begin{verbatim}
## [1] "max dim for all flakes, retouched or not"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-14.pdf}

\begin{verbatim}
## [1] "max dim for all flakes, by raw material"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-15.pdf}

\begin{verbatim}
## [1] "scar counts for all flakes, by raw material"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-16.pdf}
\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-17.pdf}

\begin{verbatim}
## [1] "facetted platforms"
\end{verbatim}

\begin{verbatim}
## 
##  fac   no 
##   43 1143
\end{verbatim}

\begin{verbatim}
## 
##   fac    no 
##  3.63 96.37
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-18.pdf}

\begin{verbatim}
## [1] "cortex proportion of flakes"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-19.pdf}

\begin{verbatim}
## [1] "compare cortex for ret and non-retouched flakes"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-20.pdf}
\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-21.pdf}

\begin{verbatim}
## [1] "platform shape vs max dim"
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-6-22.pdf}

\begin{verbatim}
## [1] "how many points?"
\end{verbatim}

\begin{verbatim}
## 
## backed  other  point 
##      9   1133     44
\end{verbatim}

\begin{verbatim}
## 
## backed  other  point 
##   0.76  95.53   3.71
\end{verbatim}

Group the flakes according to their

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Look to see if groups exist}

\CommentTok{# by flake size}
\NormalTok{plain_flakes <-}\StringTok{  }\NormalTok{flakes[}\KeywordTok{grepl}\NormalTok{(}\StringTok{"fl"}\NormalTok{, flakes}\OperatorTok{$}\NormalTok{type1), ]}
\NormalTok{d <-}\StringTok{ }\KeywordTok{density}\NormalTok{(}\KeywordTok{na.omit}\NormalTok{(plain_flakes}\OperatorTok{$}\NormalTok{mass)) }\CommentTok{# returns the density data}
\CommentTok{# plot}
\KeywordTok{plot}\NormalTok{(d)}
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-7-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(plain_flakes, }\KeywordTok{aes}\NormalTok{(mass)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_density}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-7-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# test is we can discover natural breaks that divide up the flakes?}

\CommentTok{# how many groups should we look for?}
\NormalTok{n <-}\StringTok{ }\DecValTok{5}
\NormalTok{natural_breaks <-}\StringTok{ }\KeywordTok{classIntervals}\NormalTok{(plain_flakes}\OperatorTok{$}\NormalTok{mass, }\DataTypeTok{n =}\NormalTok{ n, }
                                 \DataTypeTok{style =} \StringTok{"jenks"}\NormalTok{, }\DataTypeTok{dataPrecision =} \DecValTok{2}\NormalTok{)}
\CommentTok{# plot the breaks}
\KeywordTok{ggplot}\NormalTok{(plain_flakes, }\KeywordTok{aes}\NormalTok{(mass)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_density}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"blue"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ natural_breaks}\OperatorTok{$}\NormalTok{brks, }\DataTypeTok{colour =} \StringTok{"red"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-7-3.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# create a df}
\NormalTok{plain_flakes}\OperatorTok{$}\NormalTok{mass_group <-}\StringTok{ }\KeywordTok{cut}\NormalTok{(plain_flakes}\OperatorTok{$}\NormalTok{mass, }
                               \DataTypeTok{breaks =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, natural_breaks}\OperatorTok{$}\NormalTok{brks[}\OperatorTok{-}\DecValTok{1}\NormalTok{]),}
                               \DataTypeTok{labels =} \DecValTok{1}\OperatorTok{:}\NormalTok{(}\KeywordTok{length}\NormalTok{(natural_breaks}\OperatorTok{$}\NormalTok{brks)}\OperatorTok{-}\DecValTok{1}\NormalTok{))}
\CommentTok{# remove flakes not in a group}
\NormalTok{plain_flakes <-}\StringTok{ }\NormalTok{plain_flakes[}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(plain_flakes}\OperatorTok{$}\NormalTok{mass_group), ]}

\CommentTok{# plot the groups}
\KeywordTok{ggplot}\NormalTok{(plain_flakes, }\KeywordTok{aes}\NormalTok{(mass_group, mass)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_jitter}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-7-4.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# explore variation in thickness in these groups}
\NormalTok{thick_plot <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(plain_flakes) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_density}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\StringTok{`}\DataTypeTok{Thickness_25%max}\StringTok{`}\NormalTok{), }\DataTypeTok{colour =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_density}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\StringTok{`}\DataTypeTok{Thickness_50%max}\StringTok{`}\NormalTok{), }\DataTypeTok{colour =} \StringTok{"green"}\NormalTok{, }\DataTypeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_density}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\StringTok{`}\DataTypeTok{Thickness_75%max}\StringTok{`}\NormalTok{), }\DataTypeTok{colour =} \StringTok{"blue"}\NormalTok{, }\DataTypeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Thickness (mm)"}\NormalTok{) }

\NormalTok{thick_plot }
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-8-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# by size class that we identified earlier}
\NormalTok{thick_plot }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\StringTok{ }\NormalTok{mass_group, }\DataTypeTok{ncol =} \DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-8-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Lycett and Eren found out the thickness is more evenly distributed and less variable across preferential Levallois flakes. So , let's get the CV values for thickness for each group}
\NormalTok{CV <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(the_vector)\{}
\NormalTok{  mean_ <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(the_vector, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{  sd_ <-}\StringTok{ }\KeywordTok{sd}\NormalTok{(the_vector, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{  cv <-}\StringTok{ }\NormalTok{(sd_}\OperatorTok{/}\NormalTok{mean_)}\OperatorTok{*}\DecValTok{100}
  \KeywordTok{round}\NormalTok{(cv, }\DecValTok{3}\NormalTok{)}
\NormalTok{\}}

\NormalTok{plain_flakes }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(mass_group) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{cv_25_thick =} \KeywordTok{CV}\NormalTok{(}\StringTok{`}\DataTypeTok{Thickness_25%max}\StringTok{`}\NormalTok{),}
                   \DataTypeTok{cv_50_thick =} \KeywordTok{CV}\NormalTok{(}\StringTok{`}\DataTypeTok{Thickness_50%max}\StringTok{`}\NormalTok{),}
                   \DataTypeTok{cv_75_thick =} \KeywordTok{CV}\NormalTok{(}\StringTok{`}\DataTypeTok{Thickness_75%max}\StringTok{`}\NormalTok{),}
                   \DataTypeTok{cv_25_width =} \KeywordTok{CV}\NormalTok{(}\StringTok{`}\DataTypeTok{Width_25%max}\StringTok{`}\NormalTok{), }
                   \DataTypeTok{cv_50_width =} \KeywordTok{CV}\NormalTok{(}\StringTok{`}\DataTypeTok{Width_50%max}\StringTok{`}\NormalTok{),}
                   \DataTypeTok{cv_75_width =} \KeywordTok{CV}\NormalTok{(}\StringTok{`}\DataTypeTok{Width_75%max}\StringTok{`}\NormalTok{), }
                   \DataTypeTok{n =} \KeywordTok{n}\NormalTok{()) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 5 x 8
##   mass_group cv_25_thick cv_50_thick cv_75_thick cv_25_width cv_50_width
##   <fct>            <dbl>       <dbl>       <dbl>       <dbl>       <dbl>
## 1 1                 47.7        46.5        49.8        35.3       34.4 
## 2 2                 26.7        21.5        24.0        17.5       17.2 
## 3 3                 20.9        12.9        23.4        23.8        9.65
## 4 4                 11.1        32.1        26.4        17.7       10.6 
## 5 5                 NA          NA          NA          NA         NA   
## # ... with 2 more variables: cv_75_width <dbl>, n <int>
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# calculate the edge angle}

\NormalTok{edge_angle_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{edge_angle}

\NormalTok{edge_angle_}\DecValTok{2}\NormalTok{[, }\DecValTok{2}\OperatorTok{:}\NormalTok{(}\KeywordTok{ncol}\NormalTok{(edge_angle_}\DecValTok{2}\NormalTok{)}\OperatorTok{-}\DecValTok{1}\NormalTok{)] <-}\StringTok{ }\KeywordTok{atan}\NormalTok{(edge_angle[, }\DecValTok{2}\OperatorTok{:}\NormalTok{(}\KeywordTok{ncol}\NormalTok{(edge_angle_}\DecValTok{2}\NormalTok{)}\OperatorTok{-}\DecValTok{1}\NormalTok{)]}\OperatorTok{/}\DecValTok{2}\OperatorTok{/}\DecValTok{3}\NormalTok{)}\OperatorTok{/}\NormalTok{pi}\OperatorTok{*}\DecValTok{180}\OperatorTok{*}\DecValTok{2}

\NormalTok{df_angle <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(edge_angle_}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Using number, X__1 as id variables
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# plot the angles }
\KeywordTok{ggplot}\NormalTok{(df_angle, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ value, }\DataTypeTok{col =}\NormalTok{ variable)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_density}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{theme_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 5480 rows containing non-finite values (stat_density).
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-9-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## find quina and compare their angles.}

\NormalTok{quina_id <-}\StringTok{ }\NormalTok{flakes}\OperatorTok{$}\NormalTok{number[}\KeywordTok{which}\NormalTok{(flakes}\OperatorTok{$}\NormalTok{type3 }\OperatorTok{==}\StringTok{ "quina"}\NormalTok{)]}

\NormalTok{df_angle}\OperatorTok{$}\NormalTok{quina <-}\StringTok{ }\KeywordTok{ifelse}\NormalTok{(df_angle}\OperatorTok{$}\NormalTok{number }\OperatorTok{%in%}\StringTok{ }\NormalTok{quina_id, }\StringTok{"quina"}\NormalTok{, }\StringTok{"no"}\NormalTok{)}

\KeywordTok{ggplot}\NormalTok{(df_angle, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ value, }\DataTypeTok{col =}\NormalTok{ quina)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_density}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{theme_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 5480 rows containing non-finite values (stat_density).
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-9-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(df_angle, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ value, }\DataTypeTok{col =}\NormalTok{ variable)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_density}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{quina)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 5480 rows containing non-finite values (stat_density).
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-9-3.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## compare thickness of quina and other}

\NormalTok{df_thick <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(flakes[, }\KeywordTok{c}\NormalTok{(}\StringTok{"number"}\NormalTok{, }\StringTok{"Thickness_25%max"}\NormalTok{, }\StringTok{"Thickness_50%max"}\NormalTok{, }\StringTok{"Thickness_75%max"}\NormalTok{)])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Using number as id variables
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df_thick}\OperatorTok{$}\NormalTok{quina <-}\StringTok{ }\KeywordTok{ifelse}\NormalTok{(df_thick}\OperatorTok{$}\NormalTok{number }\OperatorTok{%in%}\StringTok{ }\NormalTok{quina_id, }\StringTok{"quina"}\NormalTok{, }\StringTok{"no"}\NormalTok{)}

\KeywordTok{ggplot}\NormalTok{(df_thick, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ value, }\DataTypeTok{col =}\NormalTok{ quina)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_density}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{variable)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 96 rows containing non-finite values (stat_density).
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-9-4.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# plot the thickness of 50% max}
\KeywordTok{ggplot}\NormalTok{(df_thick[df_thick}\OperatorTok{$}\NormalTok{variable }\OperatorTok{==}\StringTok{"Thickness_50%max"}\NormalTok{, ], }\KeywordTok{aes}\NormalTok{(quina, value)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{theme_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 32 rows containing non-finite values (stat_boxplot).
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-9-5.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## count the number of shapes of edges.}

\NormalTok{edge_shapes <-}\StringTok{ }\KeywordTok{unlist}\NormalTok{(}\KeywordTok{strsplit}\NormalTok{(retouch}\OperatorTok{$}\StringTok{`}\DataTypeTok{edge shape}\StringTok{`}\NormalTok{, }\StringTok{","}\NormalTok{))}

\KeywordTok{data.frame}\NormalTok{(}\KeywordTok{table}\NormalTok{(edge_shapes)) }\OperatorTok{%>%}\StringTok{ }
\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{reorder}\NormalTok{(edge_shapes, Freq), }\DataTypeTok{y =}\NormalTok{ Freq)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{stat =} \StringTok{"identity"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{"n"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Edge type"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{    }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{, }\DataTypeTok{vjust =} \FloatTok{0.5}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-9-6.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data.frame}\NormalTok{(}\KeywordTok{table}\NormalTok{(edge_shapes))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   edge_shapes Freq
## 1         bec    3
## 2       borer   40
## 3         ccv  274
## 4         cvx  395
## 5        dent  134
## 6         end   32
## 7       notch  106
## 8        strt  575
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## number of layers}

\KeywordTok{f}\NormalTok{(retouch, }\StringTok{'number of layers'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-9-7.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## number of edges}

\KeywordTok{f}\NormalTok{(retouch, }\StringTok{'number of edge'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-9-8.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## number of layers for each number of eadge}

\NormalTok{df_layer_edge <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(retouch[, }\DecValTok{1}\OperatorTok{:}\DecValTok{3}\NormalTok{])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Using number as id variables
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
 \KeywordTok{data.frame}\NormalTok{(}\KeywordTok{table}\NormalTok{(retouch[,}\DecValTok{2}\OperatorTok{:}\DecValTok{3}\NormalTok{])) }\OperatorTok{%>%}\StringTok{ }
\StringTok{    }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{reorder}\NormalTok{(number.of.layers, Freq), }\DataTypeTok{y =}\NormalTok{ Freq)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{stat =} \StringTok{"identity"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{"n"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{"number of layers"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{    }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{, }\DataTypeTok{vjust =} \FloatTok{0.5}\NormalTok{)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{number.of.edge)}
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-9-9.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# scar number and direction}

\NormalTok{## count how many scars from each direction}

\KeywordTok{table}\NormalTok{(}\KeywordTok{unlist}\NormalTok{(scar_dir[,}\DecValTok{2}\OperatorTok{:}\DecValTok{8}\NormalTok{]))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
##   0   1   2   3   4   5   6   7   8 
##  85 221 168 109  73  68  61  71 138
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## calculate GIUR}

\NormalTok{df_giur <-}\StringTok{ }\NormalTok{retouch[, }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{8}\NormalTok{,}\DecValTok{10}\NormalTok{,}\DecValTok{12}\NormalTok{,}\DecValTok{14}\NormalTok{,}\DecValTok{16}\NormalTok{,}\DecValTok{18}\NormalTok{,}\DecValTok{20}\NormalTok{,}\DecValTok{22}\NormalTok{)]}

\NormalTok{df_giur[, }\OperatorTok{-}\DecValTok{1}\NormalTok{] <-}\StringTok{ }\NormalTok{df_giur[, }\OperatorTok{-}\DecValTok{1}\NormalTok{] }\OperatorTok{/}\StringTok{ }\NormalTok{retouch[, }\KeywordTok{c}\NormalTok{(}\DecValTok{9}\NormalTok{,}\DecValTok{11}\NormalTok{,}\DecValTok{13}\NormalTok{,}\DecValTok{15}\NormalTok{,}\DecValTok{17}\NormalTok{,}\DecValTok{19}\NormalTok{,}\DecValTok{21}\NormalTok{,}\DecValTok{23}\NormalTok{)]}

\NormalTok{df_giur_clustered <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(df_giur, flakes_clustered[, }\KeywordTok{c}\NormalTok{(}\StringTok{"number"}\NormalTok{, }\StringTok{"cluster"}\NormalTok{)])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Joining, by = "number"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df_giur_clustered_melt <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(df_giur_clustered[,}\OperatorTok{-}\DecValTok{1}\NormalTok{], }\DataTypeTok{id.vars =} \StringTok{"cluster"}\NormalTok{)}

\NormalTok{## compare GIUR for different mass categories}

\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{na.omit}\NormalTok{(df_giur_clustered_melt[df_giur_clustered_melt}\OperatorTok{$}\NormalTok{value }\OperatorTok{<=}\DecValTok{1}\NormalTok{,]), }\KeywordTok{aes}\NormalTok{(value)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_histogram}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{cluster)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
\end{verbatim}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-10-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## calculate the median GIUR for each group}

\KeywordTok{tapply}\NormalTok{(df_giur_clustered_melt}\OperatorTok{$}\NormalTok{value, df_giur_clustered_melt}\OperatorTok{$}\NormalTok{cluster, median, }\DataTypeTok{na.rm=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##         1         2         3         4         5 
## 0.5963639 0.5433333 0.4739884 0.4400000 0.5151515
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Invasiveness index}

\NormalTok{df_invasive <-}\StringTok{ }\NormalTok{retouch[, }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{24}\OperatorTok{:}\DecValTok{39}\NormalTok{)]}

\NormalTok{df_invasive_clustered <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(df_invasive, flakes_clustered[, }\KeywordTok{c}\NormalTok{(}\StringTok{"number"}\NormalTok{, }\StringTok{"cluster"}\NormalTok{)])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Joining, by = "number"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df_invasive_clustered_melt <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(df_invasive_clustered[,}\OperatorTok{-}\DecValTok{1}\NormalTok{], }\DataTypeTok{id.vars =} \StringTok{"cluster"}\NormalTok{)}

\NormalTok{invasive_table <-}\StringTok{ }\KeywordTok{as.data.frame}\NormalTok{(}\KeywordTok{table}\NormalTok{(df_invasive_clustered_melt[, }\KeywordTok{c}\NormalTok{(}\StringTok{"cluster"}\NormalTok{, }\StringTok{"value"}\NormalTok{)]))}

\KeywordTok{ggplot}\NormalTok{(invasive_table[invasive_table}\OperatorTok{$}\NormalTok{value }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{1.0}\NormalTok{), ], }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{value, }\DataTypeTok{y=}\NormalTok{ Freq, }\DataTypeTok{x=}\NormalTok{cluster)) }\OperatorTok{+}\StringTok{ }
\StringTok{    }\KeywordTok{geom_bar}\NormalTok{( }\DataTypeTok{stat=}\StringTok{"identity"}\NormalTok{, }\DataTypeTok{position=}\StringTok{"fill"}\NormalTok{, }\DataTypeTok{width =} \FloatTok{0.6}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{14}\NormalTok{), }\DataTypeTok{axis.title=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{16}\NormalTok{,}\DataTypeTok{face=}\StringTok{"bold"}\NormalTok{)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.text=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{14}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{for_paper_ii_files/figure-latex/unnamed-chunk-10-2.pdf}

\end{document}

### 2021-06-25 README.Rmd
add renv and update readme

@@
 # Please put your title here to include it in the file below.
-Title <- "Diverse lithic production strategies in southwest China during Late Middle Pleistocene "
+Title <- "Diverse lithic production strategies in southwest China during Late Middle Pleistocene"
 ```
 
-# lithicprodswchinalatemiddlepleistocene
+# Diverse lithic production strategies in southwest China during Late Middle Pleistocene
 
@@
 
-> Yue Hu, Ben Marwick, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+> Yue Hu, Ben Marwick, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.17605/OSF.IO/7B5QD>
 
@@
 - open the `.Rproj` file in RStudio
-- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
+- run `renv::restore()` to ensure you have the packages this analysis depends on (also listed in the
 [DESCRIPTION](/DESCRIPTION) file).

### 2021-07-04 analysis/paper/supplementary-information.R
update core platform prep plot

@@ -330,8 +330,10 @@ cores <-
     platform_preparation == "plane" ~ 'plain',
     platform_preparation == "faceted" ~ 'facetted',
     platform_preparation == "cortex" ~ 'cortical',
-    platform_preparation == "dehid" ~ 'dihedral'
+    platform_preparation == "dehid" ~ 'dihedral',
+    TRUE ~ as.character(platform_preparation)
   )) %>%
+  filter(platform_preparation != "n") %>%
   drop_na(platform_preparation)
 
 # "over"    "plain"   "plane"   "faceted"  "n"    "cortex"  "dehid".
@@ -364,7 +366,6 @@ core_plat_rotat <-  table(na.omit(cores$rotations))
 
 round(prop.table(core_plat_rotat), digits = 4)*100
 
-
 ## core reduction type
 print("summary for core reduction type")
 
@@ -386,7 +387,6 @@ p_core_scar_number <- ggplot(cores, aes(scar_number)) +
         axis.text.y = element_text(size = 20),
         axis.title.x = element_text(size = 25),
         axis.title.y = element_text(size = 25))
-
 
 core_scar_table <- table(na.omit(cores$scar_number))
 

### 2024-03-13 README.Rmd
update core platform prep plot
delete unneeded figures

@@
 # Please put your title here to include it in the file below.
-Title <- "Diverse lithic production strategies in southwest China during Late Middle Pleistocene"
+Title <- "Late Middle Pleistocene lithic technology at Guanyindong cave, Southwest China: Diversity and innovation during the Chinese Middle Palaeolithic"
 ```
 
-# Diverse lithic production strategies in southwest China during Late Middle Pleistocene
-
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/lithicprodswchinalatemiddlepleistocene/master?urlpath=rstudio)
+# Late Middle Pleistocene lithic technology at Guanyindong cave, Southwest China: Diversity and innovation during the Chinese Middle Palaeolithic
 
@@
 > Yue Hu, Ben Marwick, (nd). _`r Title`_.  <https://doi.org/xxx/xxx>
-
-Our pre-print is online here:
-
-> Yue Hu, Ben Marwick, (nd). _`r Title`_. , Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
 
@@
   - [:file\_folder: paper](/analysis/paper): R Markdown source document
-    for manuscript. Includes code to reproduce the figures and tables
-    generated by the analysis. It also has a rendered version,
+    for our manuscript. It also has a rendered version,
     `paper.docx`, suitable for reading (the code is replaced by figures

### 2024-03-13 analysis/paper/paper.Rmd
rearrange files slightly
delete unneeded figures

@@
 ---
-title: "Supplementary Information for Diverse lithic production strategies in southwest China during Late Middle Pleistocene"
+title: "Late Middle Pleistocene lithic technology at Guanyindong cave, Southwest China: Diversity and innovation during the Chinese Middle Palaeolithic"
 ---
@@
 <!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->
-
-<!-- With the following code you can access and display values from the yml header above. -->
-
-Keywords: `r rmarkdown::metadata$keywords`
-
-Highlights: `r rmarkdown::metadata$highlights`
 
@@
 
-Read in the data.
+## Cores
+
+This is figure 3 in the main text
+
+```{r} 
+# make Figure 3 Statistical results of cores
+
+library(dplyr)
+library(readr)
+library(knitr)
+library(ggplot2)
+library(reshape2)
+library(readxl)
+library(fitdistrplus)
+library(tidyr)
+library(classInt)
+library(cowplot)
+library(gridExtra)
+
+# read in the data
+
+file_name <-
+  here::here("analysis/data/raw_data/artefacts-after Sam_Dec2018_quina_modified-2019.xls")
+
+flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
+cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
+debris <- as.data.frame(read_excel(file_name, sheet = "chunk&debris"))
+retouch <- as.data.frame(read_excel(file_name, sheet = "retouch"))
+core_flake_scars_1 <- as.data.frame(read_excel(file_name, sheet = "core flake scars"))
+edge_angle_1 <- as.data.frame(read_excel(file_name, sheet = "edge angle"))
+retouched_pebble_chunk <- as.data.frame(read_excel(file_name, sheet = "Pebble&chunk"))
+
+scar_dir <- as.data.frame(read_excel(file_name, sheet = 'scar direction'))
+
+core_flake_scars <- core_flake_scars_1[core_flake_scars_1$number %in% cores$number, ]
+
+edge_angle <- edge_angle_1[edge_angle_1$number %in% retouch$number, ]
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+# basic details of the dataset
+
+artefact_ids <- list(flake_ids = flakes$number,
+                     core_ids = cores$number,
+                     debris_ids = debris$number)
+
+retouch_ids <- retouch$number
+
+# how many unique artefacts?
+
+totals_of_each_type <- lapply(artefact_ids, function(i) length(i))
+
+count_unique_artefacts <- sum(unlist(totals_of_each_type))
+
+# debris pieces that are retouch pieces
+
+debris_with_retouch <- intersect(retouch_ids, artefact_ids$debris_ids)
+
+proportion_debris_retouched <- length(debris_with_retouch) / count_unique_artefacts
+
+# proportion of each type
+
+type_table <- data.frame(type = names(as.data.frame(totals_of_each_type)),
+                         count = as.numeric(as.data.frame(totals_of_each_type)),
+                         proportion = round(as.numeric(prop.table(as.data.frame(totals_of_each_type))),3))
+
+# add total row at the bottom
+
+type_table$type <- as.character(type_table$type)
+
+type_table <- rbind(type_table, c("total", count_unique_artefacts, round(sum(type_table$proportion),2)))
+
+# print pretty table
+
+# print("table of lithic types")
+
+# kable(type_table)
+
+### raw materials
+
+raw_material_table <- data.frame("category" = character(),
+                                 'basalt' = numeric(),
+                                 'chert' = numeric(),
+                                 'limestone' = numeric(),
+                                 'quartz' = numeric(),
+                                 'sandstone' = numeric(),
+                                 stringsAsFactors = FALSE)
+
+res <- flakes %>%
+  mutate(complet_flk = ifelse(type1 %in% c("flake", "blade", "KBW", "tablet", "leva", "leva?",
+                                           "DBD", "crest"), "complete flake",
+                              ifelse(type1 %in% c("flake brk", "blade brk", "LVFB", "proximal"),
+                                     "flake breaks", "other"))) %>%
+  dplyr :: count(material, complet_flk) %>%
+  arrange(complet_flk) %>%
+  group_by(complet_flk) %>%
+  spread(material, n) %>%
+  replace(., is.na(.), 0) %>%
+  rename(category = complet_flk)
+
+raw_material_table <- bind_rows(raw_material_table, res[1:2,])
+
+res <- cores %>% filter(!is.na(material)) %>%
+  dplyr :: count(material) %>%
+  spread(material, n) %>%
+  mutate(category = "cores")
+
+raw_material_table <- bind_rows(raw_material_table, res)
+
+res <- debris %>% filter(!is.na(material)) %>%
+  dplyr :: count(material) %>%
+  spread(material, n) %>%
+  mutate(category = "debris")
+
+raw_material_table <- bind_rows(raw_material_table, res)
+
+res <- flakes %>%
+  filter(type1 %in% c("ret", "ret brk", "BUR", "LVT", "ret leva?", "ret pro-lepoin")) %>%
+  dplyr:: select(material, number) %>%
+  dplyr :: count(material) %>%
+  spread(material, n) %>%
+  mutate(category = "retouched flakes and breaks")
+
+raw_material_table <- bind_rows(raw_material_table, res)
+
+res <- retouched_pebble_chunk %>%
+dplyr :: count(material) %>%
+  spread(material, n) %>%
+  mutate(category = "retouched chunks")
+
+raw_material_table <- bind_rows(raw_material_table, res)
+
+###tools raw materials
+
+res <- flakes %>%
+  mutate(type2 = replace(type2, is.na(type2),"unidentifiable" )) %>%
+  filter(type1 %in% c("ret", "ret brk", "BUR", "LVT", "ret leva?", "ret pro-lepoin")) %>%
+  mutate(tooltype = ifelse(type2 == "backed knife", "backed knife",
+                      ifelse(type2 %in% c("bec", "bec,scp"), "bec",
+                       ifelse(type2 %in% c("borer","borer,dent", "borer,dent,scp",
+                                           "borer,end,scp","borer,notch",
+                                           "borer,notch,scp","borer,scp",
+                                           "borer,scp,end","borer,strt" ),
+                              "borer",
+                        ifelse(type2 %in% c("BUR", "BUR? CORE?","BURIN", "burin"),
+                               "bur",
+                         ifelse(type2 %in% c("cp", "chopper"), "chopper",
+                         ifelse(type2 =="cleaver", "cleaver",
+                          ifelse(type2 %in% c("dent", "dent,borer,scp","dent,end","dent,notch",  "dent,scp",
+                                              "dent/NOT", "dent/POINT"  ),
+                                 "denticulate",
+                                 ifelse(type2 %in% c("end", "end,borer" ,
+                                                     "end,borer,scp" ,"end,scp" ,
+                                                     "endscp" ), "endscraper",
+                                ifelse(type2 %in% c("leva point","point" ,"POINT"
+                                                    , "point break",
+                                                    "point,broken", "point?"  ),
+                                       "point",
+                                       ifelse(type2 %in% c("moon","scp",
+                                                           "sidescp" , "sidescp-cvx"  ),
+                                              "scraper",
+                                   ifelse(type2 %in% c("NATBACK", "nature backed"),
+                                          "natural backed",
+                                          ifelse(type2 %in% c( "notcch", "notch" , "notch,borer"  , "notch,end","notch,scp" ,
+                                             "notch/DENT"),
+                                             "notch",
+                                             ifelse(type2 %in% c("TANG", "TANGED",  "TANGED POINT", "TANGED POINT /NOT",
+                                               "tanged point?", "TANGED POINT?"),
+                                               "tanged point", type2)))))))))))))) %>%
+  dplyr :: count(material, tooltype) %>%
+  spread(material, n) %>%
+  replace(., is.na(.), 0) %>%
+  rename(category = tooltype)
+
+raw_material_table <- bind_rows(raw_material_table, res)
+
+raw_material_table[is.na(raw_material_table)] <- 0 ##convert na to 0
+
+raw_material_table$total <- rowSums(raw_material_table[,-1])
+
+x <- raw_material_table
+
+prop_raw_material <- cbind(category = x[, 1], round(x[, -1]/rowSums(x[, 2:(ncol(x)-1)])*100, digits = 1))
+
+# for a table like what we have in the paper, SI Table S1, we
+# need to do this
+
+# put rows in a specific order
+row_order <- c("cores", "complete flake", "flake breaks", "debris", "retouched chunks",
+               "retouched flakes and breaks",
+               "baked knife", "bec", "borer", "bur", "chopper", "cleaver", "denticulate",
+               "endscraper", "natural backed", "notch", "point", "scraper", "tanged point", "unidentifiable")
+
+# for counts
+raw_material_table_counts_tbl <-
+raw_material_table %>%
+  mutate(category = factor(category, levels = row_order)) %>%
+  arrange(category) %>%
+  filter(!is.na(category)) %>%
+  mutate(others = quartz + sandstone) %>%
+  dplyr::select(category, chert, basalt, limestone, others, total)
+
+# for proportions
+raw_material_table_props_tbl <-
+prop_raw_material %>%
+  mutate(category = factor(category, levels = row_order)) %>%
+  arrange(category) %>%
+  filter(!is.na(category)) %>%
+  mutate(others = quartz + sandstone) %>%
+  dplyr::select(category, chert, basalt, limestone, others, total)
+
+# combine both to produce Table S1
+raw_material_table_props_and_counts_tbl <-
+tibble(
+  type = raw_material_table_counts_tbl$category,
+  chert_count = raw_material_table_counts_tbl$chert,
+  chert_perc = raw_material_table_props_tbl$chert,
+  basalt_count = raw_material_table_counts_tbl$basalt,
+  basalt_perc = raw_material_table_props_tbl$basalt,
+  limestone_count = raw_material_table_counts_tbl$limestone,
+  limestone_perc = raw_material_table_props_tbl$limestone,
+  others_count = raw_material_table_counts_tbl$others,
+  others_perc = raw_material_table_props_tbl$others,
+  total = raw_material_table_counts_tbl$total
+)
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+### cores attribute summary
+core_summary <- cores[0, c(6:17)]
+
+core_summary["mean",] <- c("mean" = apply(na.omit(cores[,c(6:17)]), 2, mean))
+
+core_summary["sd",] <- c("sd" = apply(na.omit(cores[,c(6:17)]), 2, sd))
+
+core_summary["CV",] <- core_summary["sd",]/core_summary["mean",]
+
+core_summary <- rbind(core_summary,
+                      sapply(na.omit(cores[,6:17]), function(x) quantile(x, c(0.25, 0.5, 0.75))))
+
+########################################################################
+
+# function to compute frequencies and make a plot
+f <-  function(the_data, the_column) {
+  data.frame(table(the_data[the_column])) %>%
+    ggplot(aes(x = reorder(Var1, Freq), y = Freq)) +
+    geom_bar(stat = "identity") +
+    ylab("count") +
+    xlab(the_column) +
+    theme_bw() +
+    theme(axis.text.x = element_text(size = 20, angle = 45, hjust = 1, vjust = 1),
+          axis.text.y = element_text(size = 20),
+          axis.title.x = element_text(size = 25),
+          axis.title.y = element_text(size = 25)
+    )
+}
+
+##################################################################################
+# These lines will make figure for cores
+
+# some plots of frequencies for cores
+
+## geometry
+
+p_core_geometry <- f(cores, "core_geometry") + xlab("Core geometry") + ylab("Count")
+
+core_geom_table <-  table(na.omit(cores$core_geometry))
+
+## platform number
+
+p_core_platform <- ggplot(cores, aes(platform)) +
+  geom_bar() + theme_bw() +
+  xlab("Number of platform") +
+  ylab("Count") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25))
+
+### cortex proportion of cores
+
+p_core_cortex <- ggplot(cores, aes(cortex_percentage)) +
+  geom_histogram(binwidth = 10) + theme_bw() +
+  xlab("Cortex proportion (%)") +
+  ylab("Number of cores") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25))
+
+core_cortex_table <-  table(na.omit(cores$cortex_percentage))
+
+## platform type
+
+cores <-
+  cores %>%
+  mutate(platform_preparation = case_when(
+    platform_preparation == "over" ~ 'overhang\nremoval',
+    platform_preparation == "plane" ~ 'plain',
+    platform_preparation == "faceted" ~ 'facetted',
+    platform_preparation == "cortex" ~ 'cortical',
+    platform_preparation == "dehid" ~ 'dihedral',
+    TRUE ~ as.character(platform_preparation)
+  )) %>%
+  filter(platform_preparation != "n") %>%
+  drop_na(platform_preparation)
+
+# "over"    "plain"   "plane"   "faceted"  "n"    "cortex"  "dehid".
+# 'overhang removal', 'plain' (same as plane?), 'facetted', 'cortical', 'dihedral'
+
+p_core_plat_prep <- 
+  data.frame(table(cores["platform_preparation"])) %>%
+    ggplot(aes(x = reorder(platform_preparation, Freq), y = Freq)) +
+    geom_bar(stat = "identity") +
+    theme_bw() +
+  xlab("Platform preparation") +
+  ylab("Count") +
+  coord_flip()
+
+core_plat_type <-  table(na.omit(cores$platform_preparation))
+
+## platform rotation
+
+p_core_rotation <- ggplot(cores, aes(rotations)) +
+  geom_bar() + theme_bw() +
+  xlab("Number of rotation") +
+  ylab("Number of cores") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25))
+
+core_plat_rotat <-  table(na.omit(cores$rotations))
+
+## core reduction type
+
+p_core_reduction <- f(cores, "type") + xlab("Reduction type") + ylab("Count")
+
+core_type_table <- table(na.omit(cores$type))
+
+## core scar number
+
+p_core_scar_number <- ggplot(cores, aes(scar_number)) +
+  geom_bar() + theme_bw() +
+  xlab("Scar number on cores") +
+  ylab("Count") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25))
+
+core_scar_table <- table(na.omit(cores$scar_number))
+
+## core flake scar length
+
+core_flake_scars_melt <- melt(core_flake_scars_1)
+
+colnames(core_flake_scars_melt)[3] <- "scar_length"
+
+p_core_scar_length <- ggplot(core_flake_scars_melt[core_flake_scars_melt$scar_length < 500, ],
+       aes(scar_length)) +
+  geom_density() + theme_bw() +
+  xlab("Length of scar on cores (mm)") +
+  ylab("Density") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25))
+
+# Figrur 3 summary of cores is made here
+figure_cores <-
+  cowplot::plot_grid(p_core_platform,
+                     p_core_scar_number,
+                     p_core_scar_length,
+                     p_core_cortex,
+                     p_core_plat_prep,
+                     labels = "auto",
+                     label_size = 26,
+                     ncol = 3,
+                     align = 'v',
+                     axis = 'lr',
+                     #rel_heights = c(1, 1.5),
+                     vjust = 1.5,
+                     hjust = -0.5,
+                     scale = 0.95) #
+
+ # Fig 3 Statistical results of cores
+
+save_plot(here::here("analysis/figures/fig-3-cores.png"), figure_cores,
+          base_aspect_ratio = 1,
+          base_height = 13,
+          base_width = 21)
+
+knitr::include_graphics(here::here("analysis/figures/fig-3-cores.png"))
+
+```
+
+## Flakes
+
+This is figure 9 in the main text (statistical results for flakes).
+
+```{r}
+library(readxl)
+library(tidyverse)
+
+file_name <- here::here("analysis/data/raw_data/Copy of artefacts original.xls")
+
+flakes <- read_excel(file_name, sheet = "flake basics")
+cores <- read_excel(file_name, sheet = "core basics")
+debris <- read_excel(file_name, sheet = "chunk&debris")
+retouch <- read_excel(file_name, sheet = "retouch")
+```
+
+```{r}
+# Figure 9
+
+flakes_clean <- 
+  flakes %>% 
+  mutate(mass = parse_number(mass),
+         `Length/thickness ratio` = length / `Thickness at 50% max dim`)
+
+flake_hist_mass <- 
+ggplot(flakes_clean,
+       aes(mass)) +
+  geom_histogram() +
+  theme_minimal() +
+  xlab("Mass (g)") +
+  ylab("Number of flakes")
+
+flake_dens_max <- 
+ggplot(flakes_clean,
+       aes(`max dimension`)) +
+  geom_density() +
+  theme_minimal() +
+  xlab("Maximum dimension (mm)") +
+  ylab("Density")
+
+flake_lw_ratio_median <- 
+  median(flakes_clean$`Length/thickness ratio`,
+         na.rm = TRUE)
+
+flake_lw_ratio <- 
+ggplot(flakes_clean,
+       aes(`Length/thickness ratio`)) +
+  geom_histogram() +
+  theme_minimal() +
+  geom_vline(xintercept = flake_lw_ratio_median, 
+             colour = "red") +
+  xlab("Length/thickness ratio") +
+  ylab("Number of flakes")
+
+flakes_clean_long_thick <- 
+  flakes_clean %>% 
+  dplyr::select(starts_with("Thickness")) %>% 
+  pivot_longer(everything()) %>% 
+  mutate(name = str_remove(name, "Thickness at ")) %>% 
+  mutate(name = str_remove(name, " max dim"))
+
+flake_thick_dens <- 
+ggplot(flakes_clean_long_thick,
+       aes(value,
+           colour = name)) +
+  geom_density() +
+  theme_minimal() +
+  xlab("Thickness (mm)") +
+  theme(legend.position = c(0.75, 0.5)) +
+  ylab("Density") +
+   guides(colour = guide_legend("Thickness at % of\nmaximum\ndimension"))
+
+flakes_clean_long_width <- 
+  flakes_clean %>% 
+   dplyr::select(starts_with("Width")) %>% 
+  pivot_longer(everything()) %>% 
+  mutate(name = str_remove(name, "Width at ")) %>% 
+  mutate(name = str_remove(name, " max dim"))
+
+flake_width_dens <- 
+ggplot(flakes_clean_long_width,
+       aes(value,
+           colour = name)) +
+  geom_density() +
+  theme_minimal() +
+  xlab("Width (mm)") +
+  scale_x_continuous(limits = c(0, 150)) +
+  theme(legend.position = c(0.75, 0.5)) +
+  guides(colour = guide_legend("Width at % of\nmaximum\ndimension")) +
+  ylab("Density")
+
+flake_cortex_prop <- 
+  ggplot(flakes_clean,
+       aes(`cortex percentage`)) +
+  geom_bar() +
+  theme_minimal() +
+  xlab("Cortex proportion (%)") +
+  ylab("Number of flakes")
+
+flakes_clean_facetted_plat <- 
+  flakes_clean %>% 
+  mutate(platform = case_when(
+    str_detect(platform, "fac") ~ "facetted",
+    .default = "unfacetted"
+  ))
+
+flake_dens_max_fac <- 
+ggplot(flakes_clean_facetted_plat,
+       aes(`max dimension`,
+           colour = platform)) +
+  geom_density() +
+  theme_minimal() +
+  xlab("Maximum dimension (mm)")  +
+  theme(legend.position = c(0.75, 0.5)) +
+  ylab("Density")
+
+flake_scar_count <- 
+  ggplot(flakes_clean,
+       aes(`scar number`)) +
+  geom_bar() +
+  theme_minimal() +
+  xlab("Number of scars") +
+  ylab("Number of flakes")
+
+flakes_clean_5_scars <- 
+flakes_clean %>% 
+  mutate(more_than_five_scars = ifelse(`scar number` <5, 
+                                       "less than 5",
+                                       "more than 5")) %>% 
+  drop_na(more_than_five_scars)
+
+integer_breaks <- function(n = 5, ...) {
+  fxn <- function(x) {
+    breaks <- floor(pretty(x, n, ...))
+    names(breaks) <- attr(breaks, "labels")
+    breaks
+  }
+  return(fxn)
+}
+
+flake_scar_by_mass <- 
+  ggplot(flakes_clean_5_scars,
+       aes(`scar number`, 
+           group = `scar number`,
+           mass
+           )) +
+  geom_boxplot() +
+  theme_minimal() +
+  xlab("Number of scars") +
+  ylab("Mass (g)") +
+  scale_x_continuous(breaks = integer_breaks(10)) +
+  scale_y_log10()
+
+library(cowplot)
+
+# Statistical results for flakes.
+(a-d) The counts of flakes for different mass, different length/thickness ratios (dotted line shows the median value), cortex proportion and number of dorsal scars.
+(e-g) Density distribution of flakes for different maximum dimension, thickness, and width.
+(h) Comparison of density distributions of flakes with and without faceted platforms.
+(i) Box plots showing the mass difference between flakes by scar number.
+
+figure_9 <- 
+plot_grid(
+  flake_hist_mass,    flake_lw_ratio, flake_cortex_prop, 
+  flake_scar_count,   flake_dens_max, flake_thick_dens,
+  flake_width_dens,   flake_dens_max_fac,  flake_scar_by_mass,
+  nrow = 3,
+  ncol = 3,
+  labels = "auto"
+)
+
+save_plot(here::here("analysis/figures/figure_9.png"), 
+          figure_9,
+          base_aspect_ratio = 1,
+          base_height = 9,
+          base_width = 9)
+
+knitr::include_graphics(here::here("analysis/figures/figure_9.png"))
+
+```
+
+## Comparing Levallois, Quina, and Non-Levallois flakes
 
@@
   as_tibble() %>% 
-  select(L, where(is.numeric)) %>% 
+  dplyr::select(L, where(is.numeric)) %>% 
   group_by(L) %>% 
@@
 compare_lev_non_lev_vars_tbl  %>% 
-  select(contains("cv_")) %>% 
+  dplyr::select(contains("cv_")) %>% 
   stack() %>% 
@@
 ```
-
-## Comparison of Levallois flakes vs non-Levallois flakes
 
@@
 
-compare_lev_non_lev_mass_plot
 ```
@@
 
-compare_lev_non_lev_maxdim_plot
 ```
@@
 
-compare_lev_non_lev_thick_plot
 ```
-
-## Comparison of Quina tools vs non-Quina tools
 
@@
 
-compare_qui_non_qui_angle_plot
-
 ```
@@
   xlab("")
-
-compare_qui_non_qui_thick_plot
 
@@
 
-compare_qui_non_qui_giur_plot
-
 ```
 
-Plot combining multiple panels, this is Figure 1 is the main text
+Plot combining multiple panels, this is Figure 12 is the main text
 
@@
 
-save_plot(here::here("analysis/figures/fig-1-Figures-for-Leva-and-quina.png"),
+save_plot(here::here("analysis/figures/fig-12-Figures-for-Leva-and-quina.png"),
                      figure_leva_quina,
@@
 
-figure_leva_quina
+knitr::include_graphics(here::here("analysis/figures/fig-12-Figures-for-Leva-and-quina.png"))
+```
+
+## Retouch
+
+This is figure 14 in the main text
+
+```{r}
+
+###########################################################################
+
+### Results for retouch - Figure 14
+
+file_name <-
+  here::here("analysis/data/raw_data/artefacts-after Sam_Dec2018_quina_modified-2019.xls")
+
+flakes <- as.data.frame(read_excel(file_name, sheet = "flake basics"))
+cores <- as.data.frame(read_excel(file_name, sheet = "core basics"))
+debris <- as.data.frame(read_excel(file_name, sheet = "chunk&debris"))
+retouch <- as.data.frame(read_excel(file_name, sheet = "retouch"))
+core_flake_scars_1 <- as.data.frame(read_excel(file_name, sheet = "core flake scars"))
+edge_angle_1 <- as.data.frame(read_excel(file_name, sheet = "edge angle"))
+retouched_pebble_chunk <- as.data.frame(read_excel(file_name, sheet = "Pebble&chunk"))
+
+scar_dir <- as.data.frame(read_excel(file_name, sheet = 'scar direction'))
+
+core_flake_scars <- core_flake_scars_1[core_flake_scars_1$number %in% cores$number, ]
+
+edge_angle <- edge_angle_1[edge_angle_1$number %in% retouch$number, ]
+
+# calculate the edge angle
+
+edge_angle_2 <- edge_angle
+
+edge_angle_2[, 2:(ncol(edge_angle_2)-1)] <- atan(edge_angle[, 2:(ncol(edge_angle_2)-1)]/2/3)/pi*180*2
+
+df_angle <- melt(edge_angle_2[, 1:(ncol(edge_angle_2)-1)])
+
+df_angle$variable <- substr(df_angle$variable, 1, 9)
+
+colnames(df_angle)[2:3] <- c("section", "angle")
+
+# plot the angles
+
+print("edge angles")
+
+p_edge_angle <- ggplot(df_angle,
+                       aes(x = angle,
+                           col = section)) +
+  geom_density() + theme_bw() +
+  xlab("Edge angle (\u00B0)") + ylab("Density") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25),
+        legend.text = element_text(size = 15),
+        legend.position = c(0.85, 0.72),
+        legend.title = element_blank())+ theme(plot.margin=margin(10,10,10,10))
+
+## find quina and compare their angles.
+
+quina_id <- flakes$number[which(flakes$type3 == "quina")]
+
+df_angle$quina <- ifelse(df_angle$number %in% quina_id, "quina", "non-quina")
+
+p_edge_quina <- ggplot(df_angle, aes(x = angle, col = quina)) + geom_density() +
+  theme_bw() +
+  xlab("Edge angle (\u00B0)") + ylab("Density") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25),
+        legend.text = element_text(size = 16))+
+  theme(legend.position = c(0.85, 0.85),
+        legend.title = element_blank())+ theme(plot.margin=margin(10,10,10,10))
+
+## compare thickness of quina and other
+
+df_thick <- melt(flakes[, c("number", "Thickness_25%max", "Thickness_50%max", "Thickness_75%max")])
+
+df_thick$quina <- ifelse(df_thick$number %in% quina_id, "quina", "non_quina")
+
+df_thick$label <- substring(df_thick$variable, 11, 13)
+
+p_thick_quina <- ggplot(df_thick, aes(x = value, col = quina)) + geom_density() +
+  theme_bw() +
+  facet_wrap(~label) +
+  xlab("Thickness (mm)") + ylab("Density") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25),
+        legend.text = element_text(size = 15),
+        strip.text = element_text(size = 18),
+        legend.position = c(0.9, 0.9),
+        legend.title = element_blank()) + theme(plot.margin=margin(10,10,10,10))
+
+# plot the thickness of 50% max
+
+p_thick_quina_2 <- ggplot(df_thick[df_thick$variable =="Thickness_50%max", ], aes(quina, value)) +
+  geom_boxplot() + theme_bw() +
+  xlab("Thickness at 50% max. dim. (mm)") + ylab("Density") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25))+ theme(plot.margin=margin(10,10,10,10))
+
+## count the number of shapes of edges.
+
+edge_shapes <- unlist(strsplit(retouch$`edge shape`, ","))
+
+p_edge_shape <- data.frame(table(edge_shapes)) %>%
+  ggplot(aes(x = reorder(edge_shapes, Freq), y = Freq)) +
+  geom_bar(stat = "identity") +
+  ylab("n") +
+  xlab("Edge type") +
+  theme_bw() +theme(text = element_text(size = 20)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))+
+  theme(plot.margin=margin(10,10,10,10))
+
+## number of layers
+
+p_layers <- ggplot(retouch, aes(`number of layers`)) +
+  geom_bar() + theme_bw() +
+  xlab("Number of layers") +
+  ylab("Count") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25))+ theme(plot.margin=margin(10,10,10,10))
+
+## number of edges
+
+p_edge_number <-  ggplot(retouch, aes(`number of edge`)) +
+  geom_bar() + theme_bw() +
+  xlab("Number of edges") +
+  ylab("Count") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25))+
+  theme(plot.margin=margin(10,10,10,10))
+
+## number of layers for each number of edge
+
+df_layer_edge <- melt(retouch[, 1:3])
+
+retouch$layers <- ifelse(retouch$`number of layers` < 2, "layer = 1", "layer > 1")
+
+retouch$edges <- ifelse(retouch$`number of edge` > 3, "edge no. > 3",
+                        paste("edge no. = ", retouch$`number of edge`))
+
+p_edge_layer <- data.frame(table(retouch[,c("layers", "edges")])) %>%
+  mutate(edges =  stringr::str_remove(edges, "edge no. ")) %>%
+  mutate(edges =  stringr::str_squish(stringr::str_remove(edges, "="))) %>%
+  mutate(edges =  factor(edges, levels = c("0", "1", "2", "3", "> 3"))) %>%
+  ggplot(aes(x = edges,
+             y = Freq,
+             fill = layers)) +
+  geom_bar(stat = "identity") +
+  ylab("n") +
+  xlab("Number of edges") +
+  theme_bw() +
+  theme(axis.text.x = element_text(size = 20),
+                    axis.text.y = element_text(size = 20),
+                    axis.title.x = element_text(size = 25),
+                    axis.title.y = element_text(size = 25)) +
+  theme(plot.margin=margin(10,10,10,10)) +
+  theme(strip.text = element_text(size = 18),
+        legend.position = c(0.8, 0.9),
+        legend.title = element_blank())
+
+# scar number and direction
+
+## count how many scars from each direction
+
+## How many % flakes that have dorsal scars from at least 3 complete different directions
+
+# print("How many % flakes that have dorsal scars from at least 3 complete different directions")
+
+# table(unlist(scar_dir[,2:8]))
+
+## calculate GIUR
+
+# text says ". Most specimens were extensively retouched, i.e.,
+# more than XX% have a GIUR value greater than 0.X. "
+
+# we need to compute the GIUR as T/t for each sector, then compute the mean of all sectors
+# for each artefact
+
+# looking at the GIUR data... there are some duplicate artefact numbers, which is a nuisance
+
+# make the number ID unique
+retouch$number <- make.unique(as.character(retouch$number), sep = "_")
+
+retouch_giur_per_section <-
+  retouch %>%
+  dplyr::select(grep("^number$|_t|_T", names(retouch))) %>%
+  gather(variable, value, -number) %>%
+  separate(variable, c("section", "tee"), sep = 10) %>%
+  pivot_wider(names_from =  tee, values_from = value) %>%
+  mutate(t_T = t / T)  %>%
+  filter(t_T <= 1)
+
+retouch_giur_per_artefact <-
+  retouch_giur_per_section %>%
+  group_by(number) %>%
+  summarise(mean_giur = mean(t_T, na.rm = TRUE))
+
+# mean(retouch_giur_per_artefact$mean_giur)
+
+# how many higher than some value
+# sum(retouch_giur_per_artefact$mean_giur >= 0.5) / nrow(retouch_giur_per_artefact) * 100
+
+# join metrics variables with GIUR data
+library(stringr)
+retouch_giur_per_artefact_with_metrics <-
+  left_join(retouch_giur_per_artefact,
+            flakes,
+            by = "number") %>%
+  # filter out those with repeated IDs
+  filter(!str_detect(number, "_"))
+
+# the average max dimension of retouched piece
+average_max_dimension_of_retouch <- round(mean(retouch_giur_per_artefact_with_metrics$max.dimension, na.rm = TRUE),2)
+# what percentage of retouched pieces have a GIUR greater than N
+N <-  0.5
+percentage_with_giur_greater_than_N <-
+  round(mean(retouch_giur_per_artefact$mean_giur >= N)  * 100, 0)
+
+#----------
+# df_giur <- retouch[, c(1,8,10,12,14,16,18,20,22)]
+
+# df_giur[, -1] <- df_giur[, -1] / retouch[, c(9,11,13,15,17,19,21,23)]
+
+df_giur <- retouch_giur_per_artefact_with_metrics
+
+df_giur_clustered <- left_join(df_giur, flakes_clustered[, c("number", "cluster")])
+
+df_giur_clustered_melt <- melt(df_giur_clustered[,-1], id.vars = "cluster")
+
+## compare GIUR for different mass categories
+
+# print("compare GIUR for different mass groups")
+
+df_giur_clustered_melt$label <-
+  paste("Cluster",df_giur_clustered_melt$cluster)
+
+df_giur_clustered_melt <-
+df_giur_clustered_melt %>%
+  mutate(value = parse_number(value)) %>%
+  filter(variable == "mean_giur") %>%
+  filter(!is.na(cluster))
+
+## calculate the median GIUR for each group
+
+#print("the median GIUR for each group")
+
+#tapply(df_giur_clustered_melt$value, df_giur_clustered_melt$cluster, median, na.rm=TRUE)
+
+df_giur_clustered_melt_medians <-
+df_giur_clustered_melt %>%
+  group_by(label) %>%
+  summarise(median_giur = median(value))
+
+giur_overall_median <- tibble(giur_overall_median = median(df_giur_clustered_melt$value))
+
+library(ggbeeswarm)
+p_GIUR_mass <-
+  ggplot(na.omit(df_giur_clustered_melt[df_giur_clustered_melt$value <= 1,]),
+                      aes(label, value)) +
+  geom_boxplot() +
+  geom_quasirandom(alpha = 0.2) +
+  theme_bw() +
+  ylab("GIUR") +
+  xlab("") +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25)) + theme(plot.margin=margin(10,10,10,10))
+
+# Invasiveness index
+# print("results for invasiveness index")
+
+df_invasive <- retouch[, c(1, 24:39)]
+
+# need to compute the II as sum of each section divided by the number of sections
+df_invasive_ii <-
+df_invasive %>%
+  rowwise() %>%
+  mutate(II = sum(section_1_II,
+                  section_2_II,
+                  section_3_II,
+                  is.numeric(section_4_II),
+                  section_5_II,
+                  section_6_II,
+                  section_7_II,
+                  section_8_II,
+                  section_9_II,
+                  section_10_II,
+                  section_11_II,
+                  section_12_II,
+                  section_13_II,
+                  section_14_II,
+                  section_15_II,
+                  section_16_II,
+                  na.rm = TRUE) / 16) %>%
+  dplyr::select(number, II)
+
+df_invasive_clustered <- left_join(df_invasive_ii, flakes_clustered[, c("number", "cluster")])
+
+df_invasive_clustered_melt <-
+  melt(df_invasive_clustered[,-1], id.vars = "cluster") %>%
+  filter(!is.na(cluster)) %>%
+  mutate(label = paste0("Cluster ", cluster))
+
+II_overall_median <- tibble(II_overall_median = median(df_invasive_clustered_melt$value))
+
+p_II <-
+  ggplot(df_invasive_clustered_melt) +
+         aes(as.factor(label),
+             value) +
+  geom_boxplot() +
+  geom_quasirandom(alpha = 0.1) +
+  theme_bw() +
+  xlab("") +
+  ylab("II") +
+  ylim(0, 0.5) +
+  theme(axis.text.x = element_text(size = 20),
+        axis.text.y = element_text(size = 20),
+        axis.title.x = element_text(size = 25),
+        axis.title.y = element_text(size = 25)) + theme(plot.margin=margin(10,10,10,10))
+
+### plot together
+
+figure_retouch <-
+  cowplot::plot_grid(plot_grid(
+                              #p_flake_ret_dim,
+                               p_edge_shape,
+                               p_edge_angle, ncol = 3,
+                               labels = "auto",
+                               label_size = 26,
+                               align = 'vh',
+                               axis = 'lr'),
+                     plot_grid(p_GIUR_mass,
+                               p_II,
+                               ncol = 1, labels = c("d", "e"), label_size = 26),
+                     plot_grid(
+                               p_edge_number,
+                               p_edge_layer,
+                               ncol = 2, rel_widths = c(1.5,1.5),
+                               labels = c( "f", "g"),
+                               label_size = 26,
+                               align = 'vh',
+                               axis = 'lr'),
+                     ncol = 1,
+                     align = 'vh',
+                     axis = 'lr',
+                     #rel_heights = c(1, 1.5),
+                     vjust = 1,
+                     hjust = -0.5,
+                     scale = 0.9) #
+
+save_plot(here::here("analysis/figures/Figure-14-retouch.png"), figure_retouch,
+          base_aspect_ratio = 1,
+          base_height = 21,
+          base_width = 22)
+
+knitr::include_graphics(here::here("analysis/figures/Figure-14-retouch.png"))
 ```

### 2024-03-13 analysis/paper/supplementary-information.R

@@ -420,7 +420,7 @@ figure_cores <-
                      p_core_scar_number,
                      p_core_scar_length,
                      p_core_cortex,
-                     p_core_plat_prep,
+                    # p_core_plat_prep,
                      labels = "auto",
                      label_size = 26,
                      ncol = 3,
@@ -431,7 +431,7 @@ figure_cores <-
                      hjust = -0.5,
                      scale = 0.95) #
 
-save_plot("analysis/figures/S2 figures for cores.png", figure_cores,
+save_plot("analysis/figures/fig-3-cores.png", figure_cores,
           base_aspect_ratio = 1,
           base_height = 13,
           base_width = 21)
